### 2024-06-07 manuscript.Rmd
manuscript and bib added

---
title             : "`reprodubileRchunks`: R Markdown Code Chunks with Testable Reproducibility"
shorttitle        : "Title"

author: 
  - name          : "Andreas M. Brandmaier"
    affiliation   : "1,2,3"
    corresponding : no    # Define only one corresponding author
    address       : "Rüdesheimer Str. 50, 14197 Berlin"
    email         : "andreas.brandmaier@medicalschool-berlin.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

  - name          : "Aaron Peikert"
    affiliation   : "2,3"
    corresponding : no    # Define only one corresponding author
    email         : "peikert@mpib-berlin.mpg.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, MSB Medical School Berlin"
  - id            : "2"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

The `reproducibleRchunk` package allows you to make computational results in R testable for reproduction (does the same script with the same data produce the same results, e.g. on a different computer and/or later in time) with minimal changes to your workflow. 
There is only a single thing you need to change in your analysis if you are already using RMarkdown: In the first code chunk of your document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
In `reproducibleR` chunks, all newly declared variables are identified and there contents are stored in a meta data file. 
Once a document is reproduced, in R-Markdown speak _knitted_ again [@yihui2015],
A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted document (i.e., the resulting HTML, PDF or DOCX file).
Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over which stages of a data analysis or any other computation should be checked for reproducibility. 
For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of the analysis.
In the following, we will briefly demonstrate how successful and failed reproduction will be displayed, we will go into detail of the mechanics of the package briefly and discuss some limitations of the approach.

# Demonstration

The functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents, which can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters). 
The package should be loaded in the first regular `R` code chunk in the document using
```{r eval=FALSE}
library(reproducibleRchunks)
```

![](img/rstudio-screenshot-marker2.png)

![](img/generation-step1.png)

![](img/generation-step2.png)

# Methods

- reproducibleCodechunks - what is stored
- hashing vs non-hashing
- how does the JSON look like?
- meta-data payload

The package executes reproducibleR code chunks as regular R code.
Once code execution is completed, it gathers information about all variables that were newly declared in the current chunk. 
The contents of those variables are stored in a separate JSON data file.
The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`). 
That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored
Once the document is re-generated and JSON data files exist, their content is checked against the newly computed chunk variables for identity.

```{r}
set.seed(42)
numbers_ex <- rnorm(5)
```

Here is an example of how the contents of two objects are stored, which is a single variable called numbers with a vector of five numbers `r numbers_ex` (rounded to a specified precision):

```{r}
jsonlite::serializeJSON(numbers_ex)
```
## Comparisons

![](img/schema-json-fingerprints.png)

Once a document is (re-)generated with reproducibility data present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects amenable for reproducibility checks.
Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
Using exact comparisons of object identity is sometimes more restrictive than one would expect.
This is particularly true for numeric comparisons. 
For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit). To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ...)
```{r}
jsonlite::serializeJSON(digest::digest(numbers_ex))
```

## Tutorial

In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents.

First, we define a reproducible code chunk by setting the code chunk language to `reproducibleR`. 
Once this document is rendered the first time, a data file will be created, which stores all reproducible results, which are computed in this code block. The name of the file will contain the label name (in this example `abc`). The following code block contains two reproducible results stored in variables `x` and `y`. Those computations can be based on results computed in previous chunks (here, variable `e`).
If the R markdown document is rendered a second time, the computational results are recomputed and compared against the original results. 
For each result, success or failure or reproduction is reported separately.

```{reproducibleR abc}
xu <- rnorm(10, mean=0, sd=1) + e
y <- 4 * 4
```

If nothing goes wrong, rendering this markdown twice should result in a message of successful replication. To break reproducibility, remove or comment out the `set.seed()` command and render the document again. Now, you should obtain an error message indicating that the result of `x` could not be reproduced while `y` still reproduces as it is not dependend on the random number generator.

In practice, we expect cases, in which the reproduction report should be displayed (e.g., when testing and reproducing) or not (e.g., when rendering a presentation or a manuscript for submission to a journal). By default, reproduction report statements are produced for each chunk. Displaying these statements can be surpressed on a chunk-by-chunk basis by setting the chunk option

### Changing defaults

Some default behavior of the package can be changed via R `options()`. The package defines the following options:
- reproducibleRchunks.digits. This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.

# What could possible go wrong?

In the following, we briefly discuss a few situations, in which threats to reproducibility happened and explain how the package handles these situations by default

- A markdown

- The `reproducibleRchunks` package is not available anymore

- `R` is not available anymore. As with every fashion, some day `R` will be superseded by a future programming language and/or there will be no port of `R` to a future computing platform. Then,

# Discussion

We strongly discourage making reproducibility checks on complete raw data. 
This is mainly for two reasons.
First, there may arise unwanted data protection issues when raw data is (accidentally) stored, copied or published via the reproducibility JSON files.
Second, the JSON format is inefficient.
It is designed to be human-readable and conformable with versioning tools such as git.
It is not meant for storing large data and will blow up file size by a considerable factor (while at the same time losing numerical precision).
At the same time, ensuring that the dataset identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2024-06-07 manuscript.Rmd
added images

@@
 Some default behavior of the package can be changed via R `options()`. The package defines the following options:
-- reproducibleRchunks.digits. This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
+- reproducibleRchunks.digits: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
+- reproducibleRchunks.filetype: By default this is 'json' 
+- reproducibleRchunks.hashing:
+- reproducibleRchunks.hashing_algorithm: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the `digest` package is used
+- reproducibleRchunks.templates:
 

### 2024-06-10 manuscript.Rmd
more explanations, some refs.

@@
 
-classoption       : "man"
+classoption       : "man" #man
 output            : papaja::apa6_pdf
@@
 library("papaja")
+library("reproducibleRchunks")
 r_refs("r-references.bib")
@@
 
-The `reproducibleRchunk` package allows you to make computational results in R testable for reproduction (does the same script with the same data produce the same results, e.g.
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document.
+When ever the document is compiled (or, _knitted_ in R speak), the computer code chunks are executed and allow for the dynamic generation of parts of the document.
+A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [].
+
+The `reproducibleRchunk` package allows you to make computational results in R automatically testable for reproduction (does the same script with the same data produce the same results, e.g.
 on a different computer and/or later in time) with minimal changes to your workflow.
-There is only a single thing you need to change in your analysis if you are already using RMarkdown: In the first code chunk of your document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
+There is only a single thing you need to change in your analysis if you are already using R Markdown: In the first code chunk of your document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
 In `reproducibleR` chunks, all newly declared variables are identified and there contents are stored in a meta data file. 
@@
 ```
+
+The package registers a new type of code chunk called `reproducibleR`.
+The following example shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. 
+In the chunk, a new variable `my_sum` is declared and defined to be the sum of a variable `x` plus one. 
+Variable `x` was declared in an earlier code chunk and is thus not subject to the reproducibility test of this chunk.
 
@@
 For each result, success or failure or reproduction is reported separately.
+```{r eval=TRUE,echo=FALSE}
+e<-1
+```
 
@@
 By default, reproduction report statements are produced for each chunk.
-Displaying these statements can be surpressed on a chunk-by-chunk basis by setting the chunk option
+Displaying these statements can be suppressed on a chunk-by-chunk basis by setting the chunk option
 
@@
 Some default behavior of the package can be changed via R `options()`. The package defines the following options:
+
 - reproducibleRchunks.digits: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
@@
 
+Here are a few examples how these options can be changed.
+
+First, it is possible to change the precision with which numeric results are stored.
+By default, this is up to 8 digits.
+The following line reduces the precision to only four digits after the decimal point:
+
+```{r}
+options(reproducibleRchunks.digits = 4)
+```
+
+By default, computational results are stored as fingerprints using a hash function. 
+To store data in raw format, switch off hashing with the following option:
+
+```{r}
+options(reproducibleRchunks.hashing = FALSE)
+```
+
+There are various hashing functions available, which have different features.
+Generally, hashing functions should map objects of arbitrary size to a fixed-size alphanumeric string, they should be efficient to compute and minimize the chance of collision, that is, the chance that to different objects are mapped on to the same fingerprint.
+Supported algorithm (through the digest package, @digest2022) are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
+To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint.
+
+```{r}
+options(reproducibleRchunks.hashing_algorithm = "sha256")
+```
+
+Note that these options can be changed for each chunk.
+That is, it is possible, to use fingerprints for storing results of one chunk and plain data storage for results of another chunk.
+
 # What could possible go wrong?
@@
 
-- A markdown
+- Some computations are executed and their results are stored in a JSON file as planned.
+Later on, someone changes the code in the R Markdown file, such that the results differ and a failure of reproduction is indicated.
 
-- The `reproducibleRchunks` package is not available anymore
+- `R` is not available anymore.
+As with every fashion, some day `R` will be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform.
+Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent.
 
-- `R` is not available anymore. As with every fashion, some day `R` will be superseded by a future programming language and/or there will be no port of `R` to a future computing platform. Then,
+- Error `In get_engine(options$engine) :
+  Unknown language engine 'reproducibleR' This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
+Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
 
 # Discussion
+
+In developing this package, we adhere to three major aspects of rigor in scientific software development [@brandmaier2024commentary].
+First, the package comes with a variety of formal tests (based on the `testthat` package, @testthat2011) to test correct functioning of the package.
+Second, we provide documentation in form of this manuscript and an online documentation ([https://github.com/brandmaier/reproducibleRchunks](https://github.com/brandmaier/reproducibleRchunks)). 
+Third, bug reports and feature requests can be submitted through our github project website.
 
@@
 It is not meant for storing large data and will blow up file size by a considerable factor (while at the same time losing numerical precision).
-At the same time, ensuring that the dataset identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame
+At the same time, ensuring that the data set identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame
 

### 2024-06-10 manuscript.Rmd
more writing....

@@
 
-R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document.
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
 When ever the document is compiled (or, _knitted_ in R speak), the computer code chunks are executed and allow for the dynamic generation of parts of the document.
-A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [].
+A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [@peikert2021reproducible].
 
@@
 
+### Types of objects
+
+What types of objects are suited for automated reproducibility tests? The answer is simple. In principle, any R object is suitable no matter how complex.
+For creating fingerprints, the package uses R's serialization methods, which take any R object as input.
+That is, all of the following variables (`x` of class _integer_, `y` of class _character_, and `z` of class _lm_) are valid examples for the automated reproducibility checks:
+
+```{r examples, eval=FALSE}
+  x <- 1:10
+  y <- "qr"
+  z <- lm(x~1, method=y)
+```
+
 ## Tutorial
@@
 
-```{r}
+```{r eval=FALSE, echo=TRUE}
 options(reproducibleRchunks.digits = 4)
@@
 
-```{r}
+```{r eval=FALSE, echo=TRUE}
 options(reproducibleRchunks.hashing = FALSE)
@@
 
-```{r}
+```{r eval=FALSE, echo=TRUE}
 options(reproducibleRchunks.hashing_algorithm = "sha256")
@@
 
-Note that these options can be changed for each chunk.
+Note that these options can be chosen differently for each chunk.
 That is, it is possible, to use fingerprints for storing results of one chunk and plain data storage for results of another chunk.
 
-# What could possible go wrong?
+### Customization
 
-In the following, we briefly discuss a few situations, in which threats to reproducibility happened and explain how the package handles these situations by default
+Last, users of this package can customize the appearance of the reproducibility reports. 
+This can be done by either using convenient template layouts, which can be tailored to each specific (pandoc) output format, that is, for example a template for HTML file and a different LaTeX template for PDF files.
+Or, users can obtain a summary of the reproduction status of every variable in every chunk. 
+This information is returned by function `get_reproducibility_summary()` which returns a `data.frame` with three columns. 
+The first column contains the name of the code chunk, the second column contains a variable name and the third column contains a boolean variable representing whether the reproduction was a success. 
+This information can be used to either generate custom reports, such as one general report at the very end of the document, or write reports to a separate file.
+The alternative is to just use the default reports that are appended to each code chunk output (as shown in the previous examples). 
+Here, users cannot change the content but have some degrees of freedom in styling the output.
+The option `reproducibleRchunks.templates` stores the default templates that are used for displaying the report information. 
+It is a list of key-value pairs where the key is the final pandoc output format (typically either html, pdf or docx) and the value is a string containing formatting information.
+If html output is chosen, it can contain any valid HTML/CSS code, if the output is pdf, it can contain any valid LaTeX code.
+These formatting instructions can contain two placeholders, which the package will replace with the title of the report (\${title}) and the content of the report (\${content}).
+
+```{r echo=TRUE, eval=FALSE}
+    options(reproducibleRchunks.templates = list(
+      html="<div style='border: 3px solid black; padding: 10px 10px 10px 10px; background-color: #EEEEEE;'><h5>${title}</h5>${content}</div>"))
+```
+
+For example, the following code reformats the appearance of the code report in PDF documents, such that the report is enclosed by two horizontal lines (\\hr elements), the title is displayed as a section header, there is a medium skip between title and content and content is displayed in a smaller font size:
+
+```{r echo=TRUE, eval=FALSE}
+    options(reproducibleRchunks.templates = list(
+      pdf="\\hr \\section{${title}} \\medskip \\smaller ${content} \\hr"
+    )
+```
+
+# Discussion
+
+We introduced an R package that allows for R code chunks in R Markdown documents that are automatically tested for reproducibility. 
+The output of the package is compatible with all pandoc output formats including the standard choices html, pdf, and docx. 
+Templates can be used to customize the output and summary functions allow users to generate arbitrary reproducibility reports either within the document or as separate files.
+
+### What could possible go wrong?
+
+In the following, we briefly discuss a few situations, in which threats to reproducibility happened and explain how the package handles these situations by default:
+
+- Original computations were executed and stored in the JSON data file.
+Later, a reproduction attempt is made using the exact same R Markdown file but on a different computer.
+This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility, @peikert2021reproducible].
+Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility.
+To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
 
@@
 
-# Discussion
-
+### Rigor in software development
 In developing this package, we adhere to three major aspects of rigor in scientific software development [@brandmaier2024commentary].
@@
 Third, bug reports and feature requests can be submitted through our github project website.
+
+### Storing and publishing
 
@@
 
+### Summary
+
+Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing reproducibility.
+
+Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [], we ourselves have encountered various situations, in which such approaches (including our own) failed.
+First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that run identical on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).
+On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., when Apple switched to their own processors) or because a user does not have admin privileges on a machine.
+Further some approaches rely on service providers such as the MRAN archive, which was unexpectedly terminated. 
+These different Achilles' heels at least occasionally lead to situations, in which users will try to locally reproduce historic computational results.
+With our package, the success of these reproductions becomes easily testable.
+
+We hope this package helps increasing the visibility of potential non-reproducibility issues in code and that this eventually increases quality of our research assets.
 \newpage

### 2024-06-11 manuscript.Rmd
added various text pieces and references

@@
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+  Please address correspondence to: Prof. Dr. Andreas M. Brandmaier, Rüdesheimerstr. 50, 14197 Berlin, Germany. Email: andreas.brandmaier@medicalschool-berlin.de
 
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+ Research is considered _reproducible_ if the same analysis on the same data yields the same results if performed on a different computer or on the same computer later in time.
+Reproducibility is a prerequisite for replicable, robust and transparent research.
+ Various approaches have been suggested to increase chances of reproducibility.
+ Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, presentation).
+ However, a simple way to test whether different parts of these assets are reproducible is still missing.
+ We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically store meta data about original computational results and check later reproductions of the research assets for reproduction.
+ With a minimal change to users' workflows, we hope that this approach increases transparency and trustworthiness of research assets.
 
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
+keywords          : "reproducibility, Markdown, R"
 wordcount         : "X"
@@
 When ever the document is compiled (or, _knitted_ in R speak), the computer code chunks are executed and allow for the dynamic generation of parts of the document.
-A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [@peikert2021reproducible].
+A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [@peikert2021reproducible];
+further, it is suitable to generate a variety of scientific assets such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
 
@@
 In `reproducibleR` chunks, all newly declared variables are identified and there contents are stored in a meta data file. 
-Once a document is reproduced, in R-Markdown speak _knitted_ again [@yihui2015],
+Furthermore, the content of the code chunks itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk  or changes of the computation.
+Once a document is reproduced, in R-Markdown speak _knitted_ again [@yihui2015], all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
 A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted document (i.e., the resulting HTML, PDF or DOCX file).
@@
 For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
-Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of the analysis.
+Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given data analysis (or any other computation).
 In the following, we will briefly demonstrate how successful and failed reproduction will be displayed, we will go into detail of the mechanics of the package briefly and discuss some limitations of the approach.
@@
 In the chunk, a new variable `my_sum` is declared and defined to be the sum of a variable `x` plus one. 
-Variable `x` was declared in an earlier code chunk and is thus not subject to the reproducibility test of this chunk.
+Variable `x` was declared in an earlier code chunk (not shown in the following screenshot) and is thus not subject to the reproducibility test of this chunk.
 
 ![](img/rstudio-screenshot-marker2.png)
+In this example, the R Markdown document is knitted into a HTML-report.
+On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file. 
+A reproducibility report is shown below the code chunk to indicate all variables of the given chunk whose data are stored:
 
 ![](img/generation-step1.png)
+Once the document is generated a second time and the reproducibility data exists, the computation is executed anew and its results are compared to the stored reproducibility information on a variable by variable basis.
+The reproducibility report will now contain information about the success of the individual reproduction attempts.
+In our example, the computation was reproduced successfully:
 
 ![](img/generation-step2.png)
+If the document is regenerated and the computational result is not identical to the original result, this failure of reproduction is indicated in the reproducibility report.
+In this example, if someone changed the value of `x`, variable `my_sum` would change and the following failure would be displayed:
+
+![](img/generation-step2-failed.png)
 
 # Methods
-
-- reproducibleCodechunks - what is stored
-- hashing vs non-hashing
-- how does the JSON look like?
-- meta-data payload
 
@@
 The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`). 
-That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored
+That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored in a file called _.repro_sem_analysis.Rmd_datacleaning.json_.
 Once the document is re-generated and JSON data files exist, their content is checked against the newly computed chunk variables for identity.
 
-```{r}
+```{r echo=FALSE}
 set.seed(42)
@@
 ```
+Here is an example of how the contents are stored, which is a single variable called `numbers` with a vector of five numbers `r numbers_ex` generated by the following command:
 
-Here is an example of how the contents of two objects are stored, which is a single variable called numbers with a vector of five numbers `r numbers_ex` (rounded to a specified precision):
+```{r echo=TRUE, eval=FALSE}
+numbers_ex <- rnorm(5)
+```
 
-```{r}
+This vector is serialized to a JSON file, which is an open and transparent format for storing contents that are both machine-readable but also human-readable:
+
+```{r echo=FALSE, eval=TRUE, results="asis"}
 jsonlite::serializeJSON(numbers_ex)
 ```
-## Comparisons
+
+## Comparisons of computational results
 
@@
 Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
-To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ...)
-```{r}
+To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits)
+```{r echo=FALSE, eval=TRUE, results="asis"}
 jsonlite::serializeJSON(digest::digest(numbers_ex))
@@
 
-```{r examples, eval=FALSE}
+```{r examples, eval=FALSE, echo=TRUE}
   x <- 1:10
@@
 
+If users wish to entirely suppress the default reproducibility reports, they can use the option `report=FALSE` as part of the code chunk options to suppress the display. 
+Note that reproduction is still attempted and reproducibility information is still accessible through the function `get_reproducibility_summary()`.
+The other standard chunk options can be used with `reproducibleR`-code chunks as usual. 
+In particular, `eval=FALSE` suppressed execution of the code, `echo=FALSE` suppressed that the code is shown and so forth.
+
 # Discussion
@@
 In the following, we briefly discuss a few situations, in which threats to reproducibility happened and explain how the package handles these situations by default:
+
+- Error `In get_engine(options$engine) :
+  Unknown language engine 'reproducibleR' This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
+Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
 
@@
 
-- Error `In get_engine(options$engine) :
-  Unknown language engine 'reproducibleR' This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
-Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
+### Rigor in software development
 
-### Rigor in software development
 In developing this package, we adhere to three major aspects of rigor in scientific software development [@brandmaier2024commentary]. 
@@
 
-### Storing and publishing
+### Storing and publishing reproducibility information
 

### 2024-06-11 manuscript.Rmd
added even more text

@@
 ---
-title             : "`reprodubileRchunks`: R Markdown Code Chunks with Testable Reproducibility"
+title             : "`reprodubileRchunks`: R Markdown Code Chunks with Automated Reproducibility Tests"
 shorttitle        : "Title"
@@
 abstract: |
- Research is considered _reproducible_ if the same analysis on the same data yields the same results if performed on a different computer or on the same computer later in time.
-Reproducibility is a prerequisite for replicable, robust and transparent research.
+ Computational results are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
+Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments.
  Various approaches have been suggested to increase chances of reproducibility.
@@
 keywords          : "reproducibility, Markdown, R"
-wordcount         : "X"
+wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')))`"
 
@@
 
+Computational results, including those from statistical data analyses, are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
+Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments.
+ Various approaches have been suggested to increase chances of reproducibility [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac].
+ Many approaches rely on R Markdown.
 R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
@@
 
-The `reproducibleRchunk` package allows you to make computational results in R automatically testable for reproduction (does the same script with the same data produce the same results, e.g.
+The `reproducibleRchunk` package allows users to make computational results in R Markdown documents automatically testable for reproduction (does the same script with the same data produce the same results, e.g.
 on a different computer and/or later in time) with minimal changes to your workflow.
-There is only a single thing you need to change in your analysis if you are already using R Markdown: In the first code chunk of your document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
-In `reproducibleR` chunks, all newly declared variables are identified and there contents are stored in a meta data file. 
-Furthermore, the content of the code chunks itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk  or changes of the computation.
-Once a document is reproduced, in R-Markdown speak _knitted_ again [@yihui2015], all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
+There is only a single thing users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
+In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a meta data file. 
+Furthermore, the code chunk itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
+Once a document is reproduced, in R-Markdown-speak _knitted_ again [@yihui2015], all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
+The result of the reproduction attempt is displayed in a reproducibility report for each chunk.
+Customizations of these reports are possible as we will demonstrate later.
 A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted document (i.e., the resulting HTML, PDF or DOCX file).
@@
 Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given data analysis (or any other computation).
-In the following, we will briefly demonstrate how successful and failed reproduction will be displayed, we will go into detail of the mechanics of the package briefly and discuss some limitations of the approach.
+Also, not all elements of a research asset are required to reproduce (such as the current date displayed in a presentation of results).
+In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will handle these.
 
@@
 
-The functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents, which can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters).
+The core functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters).
 The package should be loaded in the first regular `R` code chunk in the document using
-```{r eval=FALSE}
+```{r eval=FALSE, echo=TRUE}
 library(reproducibleRchunks)
 ```
-
-The package registers a new type of code chunk called `reproducibleR`.
+On load, the package registers a new type of code chunk called `reproducibleR`.
 The following example shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. 
@@
 On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file. 
-A reproducibility report is shown below the code chunk to indicate all variables of the given chunk whose data are stored:
+A reproducibility report is shown below the code chunk to indicate all variables of the given chunk whose data are stored.
+The following is a snippet from the HTML generated on the first ever creation of the document:
 
@@
 
-The package executes reproducibleR code chunks as regular R code.
-Once code execution is completed, it gathers information about all variables that were newly declared in the current chunk. 
+In the following, we will describe how the package stores, retrieves and compares reproducibility information.
+First of all, the package executes `reproducibleR` code chunks just as regular R code chunks.
+That is, there are no additional limits imposed on what can be computed and tested for reproducibility.
+Once code execution is completed, the package gathers information about all variables that were newly declared in the current chunk. 
 The contents of those variables are stored in a separate JSON data file.
+JSON is a shorthand for Java Script Object Notation and its main features are that it is an open standard for mapping complex objects to text files with high simplicity and readability for machines and humans [@lennon2009introduction].
 The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`). 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, eval=TRUE}
 set.seed(42)
-numbers_ex <- rnorm(5)
+numbers_ex <- sample(1:10, 5)
 ```
-Here is an example of how the contents are stored, which is a single variable called `numbers` with a vector of five numbers `r numbers_ex` generated by the following command:
+Here is an example of how the contents of a single variable called `numbers` is stored in JSON format.
+Here, the variable content is a vector of five random draws from the number one to ten: `r numbers_ex` generated by the following command:
 
 ```{r echo=TRUE, eval=FALSE}
-numbers_ex <- rnorm(5)
+numbers_ex <- sample(1:10, 5)
 ```
 
-This vector is serialized to a JSON file, which is an open and transparent format for storing contents that are both machine-readable but also human-readable:
+This vector can be serialized in raw format to a JSON format as follows:
 
@@
 
+For privacy reasons and to save disk space, we actually do not store the raw data by default but only fingerprints of the data.
+Fingerprints are realized via so-called one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
+Fingerprints allow for an efficient comparison of identity of results while avoid data security and privacy issues as they do not allow to reproduce the original data.
+The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint:
+
+```{r echo=FALSE, eval=TRUE, results="asis"}
+jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"))
+```
+
+In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores fingerprints of the code syntax.
+This is shown in the following schematic figure:
+![](img/schema-json-fingerprints.png)
+
 ## Comparisons of computational results
-
-![](img/schema-json-fingerprints.png)
 
@@
 To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits)
-```{r echo=FALSE, eval=TRUE, results="asis"}
-jsonlite::serializeJSON(digest::digest(numbers_ex))
-```
 
@@
 
-```{reproducibleR abc}
+```{reproducibleR abc,echo=TRUE,eval=TRUE}
 xu <- rnorm(10, mean=0, sd=1) + e
@@
     options(reproducibleRchunks.templates = list(
-      html="<div style='border: 3px solid black; padding: 10px 10px 10px 10px; background-color: #EEEEEE;'><h5>${title}</h5>${content}</div>"))
+      html="<div style='border: 3px solid black; 
+      padding: 10px 10px 10px 10px; 
+      background-color: #EEEEEE;'>
+      <h5>${title}</h5>${content}</div>"))
 ```
@@
     options(reproducibleRchunks.templates = list(
-      pdf="\\hr \\section{${title}} \\medskip \\smaller ${content} \\hr"
+      pdf="\\hr \\section{${title}} 
+      \\medskip \\smaller 
+      ${content} \\hr"
     )
@@
 
-- Error `In get_engine(options$engine) :
-  Unknown language engine 'reproducibleR' This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
+- Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'` This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
 Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
@@
 First, there may arise unwanted data protection issues when raw data is (accidentally) stored, copied or published via the reproducibility JSON files.
-Second, the JSON format is inefficient.
+Second, the JSON format is inefficient in terms of storage space required.
 It is designed to be human-readable and conformable with versioning tools such as git.
 It is not meant for storing large data and will blow up file size by a considerable factor (while at the same time losing numerical precision).
-At the same time, ensuring that the data set identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame
+At the same time, ensuring that the data set identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame`.
 
@@
 
-Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [], we ourselves have encountered various situations, in which such approaches (including our own) failed.
+Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
 First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that run identical on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).

### 2024-06-11 manuscript.Rmd
added image some formatting knitr images

@@
 
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : yes
@@
 
-![](img/rstudio-screenshot-marker2.png)
+```{r rstudio-screen1, fig.cap=""}
+knitr::include_graphics("img/rstudio-screenshot-marker2.png")
+```
+
 In this example, the R Markdown document is knitted into a HTML-report.
@@
 
-![](img/generation-step1.png)
+```{r rstudio-screen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is generated for the first time."}
+knitr::include_graphics("img/generation-step1.png")
+```
 Once the document is generated a second time and the reproducibility data exists, the computation is executed anew and its results are compared to the stored reproducibility information on a variable by variable basis.
@@
 
-![](img/generation-step2.png)
+```{r rstudio-screen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and a variable is successfully reproduced."}
+knitr::include_graphics("img/generation-step2.png")
+
+```
 If the document is regenerated and the computational result is not identical to the original result, this failure of reproduction is indicated in the reproducibility report.
@@
 
-![](img/generation-step2-failed.png)
+```{r rstudio-screen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed."}
+knitr::include_graphics("img/generation-step2-failed.png")
+
+```
 
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-jsonlite::serializeJSON(numbers_ex)
+jsonlite::serializeJSON(numbers_ex, pretty=TRUE)
 ```
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"))
+jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"),pretty=TRUE)
 ```
@@
 In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores fingerprints of the code syntax.
-This is shown in the following schematic figure:
-![](img/schema-json-fingerprints.png)
+This is shown in the schematic in Figure \@ref(ref:jsonschema).
+```{r jsonschema, fig.cap="(ref:jsonschema)A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax."}
+knitr::include_graphics("img/schema-json-fingerprints.png")
+
+```
 

### 2024-06-11 manuscript.Rmd
some json formatting

@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-jsonlite::serializeJSON(numbers_ex, pretty=TRUE)
+content <- jsonlite::serializeJSON(numbers_ex, pretty=TRUE)
+content <- gsub(pattern="\\n",replacement="  \n", x=content)
+cat(content)
 ```
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"),pretty=TRUE)
+content <- jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"),pretty=TRUE)
+content <- gsub(pattern="\\n",replacement="  \n", x=content)
+cat(content)
+```
+
+```{r jsonschema, fig.cap="(ref:jsonschema) A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax."}
+knitr::include_graphics("img/schema-json-fingerprints.png")
+
 ```
@@
 In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores fingerprints of the code syntax.
-This is shown in the schematic in Figure \@ref(ref:jsonschema).
-```{r jsonschema, fig.cap="(ref:jsonschema)A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax."}
-knitr::include_graphics("img/schema-json-fingerprints.png")
-
-```
+This is shown in the schematic in Figure \@ref(jsonschema).
 

### 2024-06-14 manuscript.Rmd
revisions

@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
+                      out.width = "80%")
 ```
@@
 Computational results, including those from statistical data analyses, are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
-Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments.
- Various approaches have been suggested to increase chances of reproducibility [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac].
- Many approaches rely on R Markdown.
+Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments [@epskamp2019reproducibility].
+ Various approaches have been suggested to increase chances of reproducibility [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac].
+ Many approaches rely on R Markdown [@rmarkdown2023].
 R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
-When ever the document is compiled (or, _knitted_ in R speak), the computer code chunks are executed and allow for the dynamic generation of parts of the document.
-A particular advantage of R Markdown is that it helps avoiding common threats to reproducibility of code, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a Word document) [@peikert2021reproducible];
-further, it is suitable to generate a variety of scientific assets such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
+When ever the document is compiled (or, _knitted_ in R speak), all of its computer code chunks are executed and thus generate dynamic content in the document.
+This is particularly interesting for psychological research reports, as the `papaja` package [@R-papaja] allows R Markdown documents (including tables and figures) to be automatically formatted in APA style; most importantly, statistical results and content of tables and figures are dynamically generated.
+Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
+A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a statistic or table in a Word document) [@peikert2021reproducible];
+further, it is suitable to generate a variety of scientific assets other than manuscripts such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
 
@@
 For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
-Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given data analysis (or any other computation).
+Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given statistical data analysis (or any other computation).
 Also, not all elements of a research asset are required to reproduce (such as the current date displayed in a presentation of results).
-In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will handle these.
+In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
 
@@
 The core functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters).
-The package should be loaded in the first regular `R` code chunk in the document using
+To use these new code chunks, the package must be loaded in the first regular `R` code chunk in the document using
+
 ```{r eval=FALSE, echo=TRUE}
@@
 ```
+
 On load, the package registers a new type of code chunk called `reproducibleR`.
-The following example shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. 
-In the chunk, a new variable `my_sum` is declared and defined to be the sum of a variable `x` plus one. 
-Variable `x` was declared in an earlier code chunk (not shown in the following screenshot) and is thus not subject to the reproducibility test of this chunk.
+Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. 
+In the chunk, a new variable `my_sum` is declared and defined to be the sum of some variable `x` plus one. 
+Variable `x` was declared in an earlier code chunk (not shown in the figure) and is thus not subject to the reproducibility test of this chunk.
 
-```{r rstudio-screen1, fig.cap=""}
+```{r rstudioscreen1, fig.cap="Reproducibility report shown when a document is created for the first time and no reproducibility information is available yet."}
 knitr::include_graphics("img/rstudio-screenshot-marker2.png")
@@
 
-In this example, the R Markdown document is knitted into a HTML-report.
-On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file. 
-A reproducibility report is shown below the code chunk to indicate all variables of the given chunk whose data are stored.
-The following is a snippet from the HTML generated on the first ever creation of the document:
-
-```{r rstudio-screen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is generated for the first time."}
+```{r rstudioscreen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is generated for the first time."}
 knitr::include_graphics("img/generation-step1.png")
 ```
+
+In this example, we assume that the R Markdown document is eventually knitted into a HTML-report (but all other pandoc-supported formats are supported as well).
+On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file, which contains all information for further reproducibility tests. 
+A reproducibility report is shown in the generated document below the code chunk indicating all variables of the given chunk whose data are stored.
+Figure \@ref(fig:rstudioscreen2) show a snippet from the HTML report generated on the first ever creation of the document.
+
 Once the document is generated a second time and the reproducibility data exists, the computation is executed anew and its results are compared to the stored reproducibility information on a variable by variable basis.
 The reproducibility report will now contain information about the success of the individual reproduction attempts.
-In our example, the computation was reproduced successfully:
+In our example, the computation was reproduced successfully.
+A snippet from the resulting generated document is shown in Figure \@ref(fig:rstudioscreen3).
 
-```{r rstudio-screen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and a variable is successfully reproduced."}
+```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and a variable is successfully reproduced."}
 knitr::include_graphics("img/generation-step2.png")
@@
 If the document is regenerated and the computational result is not identical to the original result, this failure of reproduction is indicated in the reproducibility report.
-In this example, if someone changed the value of `x`, variable `my_sum` would change and the following failure would be displayed:
+In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
 
-```{r rstudio-screen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed."}
+```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed."}
 knitr::include_graphics("img/generation-step2-failed.png")
@@
 Fingerprints allow for an efficient comparison of identity of results while avoid data security and privacy issues as they do not allow to reproduce the original data.
-The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint:
+The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint (which is a 256 bit-long fingerprint that is usually displayed using 64 hexadecimal characters):
 
@@
 
-```{r jsonschema, fig.cap="(ref:jsonschema) A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax."}
-knitr::include_graphics("img/schema-json-fingerprints.png")
+```{r jsonschema, fig.cap="A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax.
+This image by Andreas Brandmaier and Anke Schneider is licensed under CC BY 4.0"}
+knitr::include_graphics("img/schema-json-fingerprints_v4.png")
 
@@
 
-In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores fingerprints of the code syntax.
-This is shown in the schematic in Figure \@ref(jsonschema).
+In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores a fingerprint of the code syntax.
+This allows later checks whether a potential non-reproduction is due to a modification of the code chunk since the original computation.
+The storage of the fingerprinted information about a code chunk and its computational results is illustrated in the schematic in Figure \@ref(fig:jsonschema).
 
@@
 For creating fingerprints, the package uses R's serialization methods, which take any R object as input.
-That is, all of the following variables (`x` of class _integer_, `y` of class _character_, and `z` of class _lm_) are valid examples for the automated reproducibility checks:
+That is, all of the following variables (`x` of class _integer_, `y` of class _character_, `z` of class _lm_, and `lst` of class _list_) are valid examples for the automated reproducibility checks:
 
@@
   z <- lm(x~1, method=y)
+  lst <- list(x, y, z)
 ```
+
+Yet, we recommend subjecting simple objects to reproducibility checks, that is, preferably numeric variables such as single numbers or vectors of numbers. 
+In a typical statistical data analysis this would correspond to, for example, parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
+While virtually any R object is serializable in principle, we expect that the serialization algorithm is more likely to change over time for more complex objects in future than for simple objects.
 
@@
 
-In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents.
+In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents. First, to install the package, please install the latest version from our repository:
 
-First, we define a reproducible code chunk by setting the code chunk language to `reproducibleR`. 
+```{r eval=FALSE, echo=TRUE}
+devtools::install_github("brandmaier/reproducibleRchunks")
+```
+
+In your R Markdown file, load the package using  `library(reproducibleRchunks)`, preferably in the first regular R code chunk at the very top of the document.
+
+Now, we define a reproducible code chunk by setting the code chunk language to `reproducibleR`. 
 Once this document is rendered the first time, a data file will be created, which stores all reproducible results, which are computed in this code block. 
-The name of the file will contain the label name (in this example `abc`). 
+The name of the file will contain the label name (in this example `helloworld`). 
 The following code block contains two reproducible results stored in variables `x` and `y`. 
-Those computations can be based on results computed in previous chunks (here, variable `e`).
+Those computations can be based on results computed in previous chunks but those variables will not per se be subject to the reproducibility tests.
 If the R markdown document is rendered a second time, the computational results are recomputed and compared against the original results. 
 For each result, success or failure or reproduction is reported separately.
-```{r eval=TRUE,echo=FALSE}
-e<-1
+
+<!--
+```{reproducibleR helloworld, echo=TRUE, eval=TRUE}
+set.seed(42)
+x <- rnorm(10, mean=0, sd=1) 
+y <- 4 * 4
+```
+-->
+```{r fig.cap="A reproducibleR chunk with two variables automatically tested for reproducibility."}
+knitr::include_graphics("img/helloworld.png")
 ```
 
-```{reproducibleR abc,echo=TRUE,eval=TRUE}
-xu <- rnorm(10, mean=0, sd=1) + e
-y <- 4 * 4
+Rendering this markdown twice in a row on the same computer should result in a message of successful replication. 
+To break reproducibility, remove or comment out the `set.seed()` command, which sets the random number generator in a reproducible state, and render the document again. 
+Now, you should obtain an error message indicating that the result of `x` could not be reproduced while `y` still reproduces as it did not depend on the random number generator.
+
+In practice, we expect cases, in which the reproduction report should be displayed (e.g., when developing a data analysis script or reproducing a former one) or not (e.g., when rendering a presentation or a manuscript for submission to a journal).
+By default, reproduction report statements are produced for each chunk but this default can be changed via the code chunk argument `report=FALSE`.
+If users wish to change the default display of reports over the entire document, this can be adjusted via a global knitr option as follows:
+
+```{r echo=TRUE, eval=FALSE}
+knitr::opts_chunk$set(report = FALSE)
 ```
-
-If nothing goes wrong, rendering this markdown twice should result in a message of successful replication. 
-To break reproducibility, remove or comment out the `set.seed()` command and render the document again. 
-Now, you should obtain an error message indicating that the result of `x` could not be reproduced while `y` still reproduces as it is not dependend on the random number generator.
-
-In practice, we expect cases, in which the reproduction report should be displayed (e.g., when testing and reproducing) or not (e.g., when rendering a presentation or a manuscript for submission to a journal).
-By default, reproduction report statements are produced for each chunk. 
-Displaying these statements can be suppressed on a chunk-by-chunk basis by setting the chunk option
 
@@
 
-Some default behavior of the package can be changed via R `options()`. The package defines the following options:
+Some default behaviors of the package can be changed via R `options()`. The package defines the following options:
 
 - reproducibleRchunks.digits: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
-- reproducibleRchunks.filetype: By default this is 'json' 
-- reproducibleRchunks.hashing:
-- reproducibleRchunks.hashing_algorithm: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the `digest` package is used
-- reproducibleRchunks.templates:
+- reproducibleRchunks.filetype: This is the type of data storage. Currently, we only support the 'json' format; however, this leaves room for other formats to be supported in future. 
+- reproducibleRchunks.hashing: Boolean. Indicates whether fingerprints should be used (default) or raw values should be stored.
+- reproducibleRchunks.hashing_algorithm: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithm offered by the `digest` package are used.
+- reproducibleRchunks.templates: A list with keys corresponding to pandoc output formats and values to templates for formatting the reproducibility reports. See below for more details.
 
 Here are a few examples how these options can be changed.
-
 First, it is possible to change the precision with which numeric results are stored.
@@
 
-If users wish to entirely suppress the default reproducibility reports, they can use the option `report=FALSE` as part of the code chunk options to suppress the display. 
+As mentioned above, if users wish to entirely suppress the default reproducibility reports, they can use the chunk argument `report=FALSE` as part of the code chunk options to suppress the display. 
 Note that reproduction is still attempted and reproducibility information is still accessible through the function `get_reproducibility_summary()`.
@@
 Later, a reproduction attempt is made using the exact same R Markdown file but on a different computer.
-This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility, @peikert2021reproducible].
+This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021reproducible].
 Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility.
-To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
+To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
 
@@
 Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent.
+
+- A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken).
+The version of the serialize protocol can be manually overwritten to an older version if newer versions should not be backward compatible.
+As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2024) using this command: `options(serializeVersion=2)`.
+
+<!-- - The results of hash functions are generally not guaranteed to be stable over time. To trace back possible problems, -->
 
@@
 Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing reproducibility.
-
-Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible, @renv2022, @chan2023rang, @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
+Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
 First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that run identical on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).

### 2024-06-19 manuscript.Rmd
added default prefix info
add links to sources

@@
 authornote: |
-  Please address correspondence to: Prof. Dr. Andreas M. Brandmaier, Rüdesheimerstr. 50, 14197 Berlin, Germany. Email: andreas.brandmaier@medicalschool-berlin.de
+  Please address correspondence to: Prof. Dr. Andreas M. Brandmaier, Rüdesheimerstr. 50, 14197 Berlin, Germany. Email: andreas.brandmaier@medicalschool-berlin.de.
+  The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks.
+The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.
 
@@
 - reproducibleRchunks.templates: A list with keys corresponding to pandoc output formats and values to templates for formatting the reproducibility reports. See below for more details.
+- reproducibleRchunks.prefix: This is the prefix for the file containing reproducibility information.
+The filename will always contain the name of the R Markdown file and the chunk name separated by an underscore.
+The default prefix is `.repro`.
 

### 2024-06-19 manuscript.Rmd
Update manuscript.Rmd
In this order I find it more coherent
Update manuscript.Rmd
typo? Did not find the other part of the parenthesis
Update manuscript.Rmd
I somehow stumbled over this sentence, not completely sure why. I feel it sounds better with plural instead of singular, as "comparison" somehow implies that there need to be more than one object, but maybe this is just me

@@
 A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted document (i.e., the resulting HTML, PDF or DOCX file).
+For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over which stages of a data analysis or any other computation should be checked for reproducibility.
-For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
 Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given statistical data analysis (or any other computation).
@@
 JSON is a shorthand for Java Script Object Notation and its main features are that it is an open standard for mapping complex objects to text files with high simplicity and readability for machines and humans [@lennon2009introduction].
-The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`). 
+The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`. 
 That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored in a file called _.repro_sem_analysis.Rmd_datacleaning.json_.
@@
 Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
-Using exact comparisons of object identity is sometimes more restrictive than one would expect.
+Using exact comparisons of object identities is sometimes more restrictive than one would expect.
 This is particularly true for numeric comparisons. 

### 2024-06-19 manuscript.Rmd
added verbatim

--- 2024-06-19 manuscript.Rmd
@@
 output            : papaja::apa6_pdf
+header-includes:
+   - \usepackage{amsmath}  # providing non-breaking mbox environment
 ---
@@
 
+\begin{verbatim}
 ```{r echo=FALSE, eval=TRUE, results="asis"}
@@
 ```
+\end{verbatim}
 
@@
 ```
+\end{verbatim}
+
+In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores a fingerprint of the code syntax.
+This allows later checks whether a potential non-reproduction is due to a modification of the code chunk since the original computation.
+The storage of the fingerprinted information about a code chunk and its computational results is illustrated in the schematic in Figure \@ref(fig:jsonschema).
 
@@
 ```
-
-In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores a fingerprint of the code syntax.
-This allows later checks whether a potential non-reproduction is due to a modification of the code chunk since the original computation.
-The storage of the fingerprinted information about a code chunk and its computational results is illustrated in the schematic in Figure \@ref(fig:jsonschema).
 

### 2024-06-20 manuscript.Rmd
some minor fixes added thanks to Leonie

@@
 The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.
+We thank Leonie Hagitte for providing comments on an earlier version of the manuscript.
 
@@
 Customizations of these reports are possible as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted document (i.e., the resulting HTML, PDF or DOCX file).
+A simple way to assess reproducibility of computational results would be  a bitwise comparison of the entire knitted Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over which stages of a data analysis or any other computation should be checked for reproducibility.
@@
 Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given statistical data analysis (or any other computation).
-Also, not all elements of a research asset are required to reproduce (such as the current date displayed in a presentation of results).
+Also, not all elements of a research asset may be required to reproduce (such as the current date displayed in a presentation of results).
 In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
@@
 JSON is a shorthand for Java Script Object Notation and its main features are that it is an open standard for mapping complex objects to text files with high simplicity and readability for machines and humans [@lennon2009introduction].
-The name of the JSON file contains the original Markdown file and the chunk label and always starts with the prefix `repro`). 
+The name of the JSON file contains the original Markdown file and the chunk label and, by package default, starts with the prefix `.repro`. 
 That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored in a file called _.repro_sem_analysis.Rmd_datacleaning.json_.
@@
 set.seed(42)
-numbers_ex <- sample(1:10, 5)
+numbers <- sample(1:10, 5)
 ```
 Here is an example of how the contents of a single variable called `numbers` is stored in JSON format.
-Here, the variable content is a vector of five random draws from the number one to ten: `r numbers_ex` generated by the following command:
+Here, the variable content is a vector of five random draws from the number one to ten: `r numbers` generated by the following command:
 
 ```{r echo=TRUE, eval=FALSE}
-numbers_ex <- sample(1:10, 5)
+set.seed(42)
+numbers <- sample(1:10, 5)
 ```
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-content <- jsonlite::serializeJSON(numbers_ex, pretty=TRUE)
+content <- jsonlite::serializeJSON(numbers, pretty=TRUE)
 content <- gsub(pattern="\\n",replacement="  \n", x=content)
@@
 
+The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans. 
+Without going into detail, we believe it suffices to say that for each object, its type (character, integer, list, class, etc.), its value and its attributes are stored whereas values itself can again be collections of objects; in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
 For privacy reasons and to save disk space, we actually do not store the raw data by default but only fingerprints of the data.
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-content <- jsonlite::serializeJSON(digest::digest(numbers_ex,algo = "sha256"),pretty=TRUE)
+content <- jsonlite::serializeJSON(digest::digest(numbers,algo = "sha256"),pretty=TRUE)
 content <- gsub(pattern="\\n",replacement="  \n", x=content)
@@
 Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
-Using exact comparisons of object identity is sometimes more restrictive than one would expect.
+Using exact comparisons for testing object identity is sometimes more restrictive than one would expect.
 This is particularly true for numeric comparisons. 

### 2024-06-20 manuscript.Rmd
rewrite in response to LH

@@
 When ever the document is compiled (or, _knitted_ in R speak), all of its computer code chunks are executed and thus generate dynamic content in the document.
-This is particularly interesting for psychological research reports, as the `papaja` package [@R-papaja] allows R Markdown documents (including tables and figures) to be automatically formatted in APA style; most importantly, statistical results and content of tables and figures are dynamically generated.
+This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated. 
+Packages like `papaja` package [@R-papaja] further extend R Markdown's capabilities to format documents (including tables and figures) in a specific journal style, such as APA style.
 Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.

### 2024-06-21 manuscript.Rmd
a couple of uncontroversial changes

@@
 ---
-title             : "`reprodubileRchunks`: R Markdown Code Chunks with Automated Reproducibility Tests"
+title             : "reprodubileRchunks: R Markdown Code Chunks with Automated Reproducibility Tests"
 shorttitle        : "Title"
-
 author: 
@@
  Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, presentation).
- However, a simple way to test whether different parts of these assets are reproducible is still missing.
+ However, a simple way detect changes of these assets, i.e. irreproducibility, is still missing.
  We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically store meta data about original computational results and check later reproductions of the research assets for reproduction.
@@
 R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
-When ever the document is compiled (or, _knitted_ in R speak), all of its computer code chunks are executed and thus generate dynamic content in the document.
+When ever the document is reproduced, all of its computer code chunks are executed and thus generate dynamic content in the document.
 This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated. 
@@
 on a different computer and/or later in time) with minimal changes to your workflow.
-There is only a single thing users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
+There are only two things users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a meta data file. 
 Furthermore, the code chunk itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
-Once a document is reproduced, in R-Markdown-speak _knitted_ again [@yihui2015], all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
+Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
 The result of the reproduction attempt is displayed in a reproducibility report for each chunk.
@@
 Also, not all elements of a research asset may be required to reproduce (such as the current date displayed in a presentation of results).
+This enables more detailed forensics in case of irreproducibility.
 In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
@@
 ```
+
 If the document is regenerated and the computational result is not identical to the original result, this failure of reproduction is indicated in the reproducibility report.
@@
 ```
+
 Here is an example of how the contents of a single variable called `numbers` is stored in JSON format.
@@
 ```{r jsonschema, fig.cap="A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax.
-This image by Andreas Brandmaier and Anke Schneider is licensed under CC BY 4.0"}
+This image by Andreas Brandmaier and Anke Schneider is licensed under [CC BY 4.0]( https://creativecommons.org/licenses/by/4.0/)"}
 knitr::include_graphics("img/schema-json-fingerprints_v4.png")
-
 ```
@@
 Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
-Using exact comparisons for testing object identity is sometimes more restrictive than one would expect.
-This is particularly true for numeric comparisons. 
-For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
-Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
-To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits)
 
@@
 While virtually any R object is serializable in principle, we expect that the serialization algorithm is more likely to change over time for more complex objects in future than for simple objects.
+Using exact comparisons for testing compley object identity is sometimes more restrictive than one would expect.
+This is particularly true for numeric comparisons. 
+For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
+Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
+To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
 
@@
 ```{r eval=FALSE, echo=TRUE}
-devtools::install_github("brandmaier/reproducibleRchunks")
+if(!requireNamespace("pak"))install.packages("pak")
+pak::pkg_install("brandmaier/reproducibleRchunks")
 ```
@@
 
-<!--
-```{reproducibleR helloworld, echo=TRUE, eval=TRUE}
+````markdown
+`r ''````{reproducibleR helloworld, echo=TRUE, eval=TRUE}
 set.seed(42)
@@
 ```
--->
-```{r fig.cap="A reproducibleR chunk with two variables automatically tested for reproducibility."}
-knitr::include_graphics("img/helloworld.png")
-```
+````
 
@@
 
-- reproducibleRchunks.digits: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
-- reproducibleRchunks.filetype: This is the type of data storage. Currently, we only support the 'json' format; however, this leaves room for other formats to be supported in future. 
-- reproducibleRchunks.hashing: Boolean. Indicates whether fingerprints should be used (default) or raw values should be stored.
-- reproducibleRchunks.hashing_algorithm: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithm offered by the `digest` package are used.
-- reproducibleRchunks.templates: A list with keys corresponding to pandoc output formats and values to templates for formatting the reproducibility reports. See below for more details.
-- reproducibleRchunks.prefix: This is the prefix for the file containing reproducibility information.
+`reproducibleRchunks.digits`
+: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
+
+`reproducibleRchunks.filetype`
+: This is the type of data storage. Currently, we only support the 'json' format; however, this leaves room for other formats to be supported in future.
+
+`reproducibleRchunks.hashing`
+: Boolean. Indicates whether fingerprints should be used (default) or raw values should be stored.
+
+`reproducibleRchunks.hashing_algorithm`
+: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithm offered by the `digest` package are used.
+
+`reproducibleRchunks.templates`
+: A list with keys corresponding to pandoc output formats and values to templates for formatting the reproducibility reports. See below for more details.
+
+`reproducibleRchunks.prefix`
+: This is the prefix for the file containing reproducibility information.
 The filename will always contain the name of the R Markdown file and the chunk name separated by an underscore.
@@
 
+```{r}
+# sha benchmark
+set.seed(345)
+benchmark <- microbenchmark::microbenchmark(digest::sha1(numbers), unit = "seconds", check = "equal", setup = {numbers <- runif(1000)})
+```
+
 There are various hashing functions available, which have different features.
-Generally, hashing functions should map objects of arbitrary size to a fixed-size alphanumeric string, they should be efficient to compute and minimize the chance of collision, that is, the chance that to different objects are mapped on to the same fingerprint.
+They generally differ along three dimensions: speed, chance of collisions, and security.
+Generally, hashing functions should map objects of arbitrary size to a fixed-size alphanumeric string.
+In this application, the choice of algorithm is not overly crucial.
 Supported algorithm (through the digest package, @digest2022) are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
-To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint.
+To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of 100 numbers takes `r mean(benchmark$time)` nanoseconds).
 

### 2024-06-21 manuscript.Rmd
reinstate backtics in title
typo

@@
 ---
-title             : "reprodubileRchunks: R Markdown Code Chunks with Automated Reproducibility Tests"
+title             : "`reprodubileRchunks`: R Markdown Code Chunks with Automated Reproducibility Tests"
 shorttitle        : "Title"
@@
  Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, presentation).
- However, a simple way detect changes of these assets, i.e. irreproducibility, is still missing.
+ However, a simple way to detect changes of these assets, i.e. irreproducibility, is still missing.
  We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically store meta data about original computational results and check later reproductions of the research assets for reproduction.

### 2024-06-21 manuscript.Rmd
added stargazer reference made simulation reporting consistent

@@
 This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated. 
-Packages like `papaja` package [@R-papaja] further extend R Markdown's capabilities to format documents (including tables and figures) in a specific journal style, such as APA style.
+Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables and figures) in a specific journal style, such as APA style.
 Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
@@
 set.seed(345)
-benchmark <- microbenchmark::microbenchmark(digest::sha1(numbers), unit = "seconds", check = "equal", setup = {numbers <- runif(1000)})
+numbers_in_simulation <- 1000
+benchmark <- microbenchmark::microbenchmark(digest::sha1(numbers), unit = "seconds", check = "equal", setup = {numbers <- runif(numbers_in_simulation)})
+
+mean_time_s <- summary(benchmark)$mean
 ```
@@
 Supported algorithm (through the digest package, @digest2022) are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
-To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of 100 numbers takes `r mean(benchmark$time)` nanoseconds).
+To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers takes `r mean(benchmark$time)` seconds on our standard computers).
+We believe that in most cases, the speed factor will be irrelevant but it could become an issue when very large objects (e.g., entire neuroimaging data sets) are fingerprinted.
+In that case, users either have to use a faster fingerprinting algorithm or forego fingerprinting raw data and instead fingerprint intermediate or final results only.
 

### 2024-06-24 manuscript.Rmd
shortened title

@@
 ---
-title             : "`reprodubileRchunks`: R Markdown Code Chunks with Automated Reproducibility Tests"
-shorttitle        : "Title"
+title             : "Automated Reproducibility Testing in R Markdown Code Chunks"
+shorttitle        : "Reproducible R Chunks"
 author: 
@@
 Computational results, including those from statistical data analyses, are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
-Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments [@epskamp2019reproducibility].
- Various approaches have been suggested to increase chances of reproducibility [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac].
+Reproducibility is a prerequisite for replicable, robust and credible research in digital environments [@epskamp2019reproducibility; @hardwicke2018data].
+ Various approaches have been suggested to increase chances of reproducibility of statistical data analyses in the R language [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac].
  Many approaches rely on R Markdown [@rmarkdown2023].
@@
 Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
-First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that run identical on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).
+First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that execute code under identical conditions on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).
 On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., when Apple switched to their own processors) or because a user does not have admin privileges on a machine.
-Further some approaches rely on service providers such as the MRAN archive, which was unexpectedly terminated. 
-These different Achilles' heels at least occasionally lead to situations, in which users will try to locally reproduce historic computational results.
-With our package, the success of these reproductions becomes easily testable.
+Further some approaches rely on service providers such as the MRAN archive, which a while ago was unexpectedly terminated. 
+These different Achilles' heels now and then lead to situations, in which users will try to locally reproduce historic computational results and they may want to ensure that these reproductions were succesful.
+With our package, the success of such reproduction attempts becomes easily testable.
 
-We hope this package helps increasing the visibility of potential non-reproducibility issues in code and that this eventually increases quality of our research assets.
+We hope the suggested approach helps raising awareness for the importance of reproduciblity and increasing the visibility of potential non-reproducibility issues in R code, eventually increasing quality and credibility of digital research assets.
 \newpage

### 2024-06-25 manuscript.Rmd
more concrete advice

@@
 ---
-title             : "Automated Reproducibility Testing in R Markdown Code Chunks"
-shorttitle        : "Reproducible R Chunks"
+title             : "Automated Reproducibility Testing in R Markdown"
+shorttitle        : "Reproducibility Testing"
 author: 
@@
 
-# Demonstration
+# Introductory Example
 
@@
 
+### Recommendations
+
+In the following, we give recommendations on how to use the package in typical cases of statistical data analysis and reporting.
+It is possible to simply replace all classic `R` code chunks with `reproducibleR` code chunks.
+Then, all variables created in the process of a given data analysis will be tested for reproducibility.
+However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results will perfectly reproduce.
+
+```{r eval=TRUE, include=FALSE}
+library(OpenMx)
+
+data(demoOneFactor)
+manifests <- names(demoOneFactor)
+latents   <- c("G")
+
+factorModel1 <- mxModel(name="One Factor",
+                        type="RAM",
+                        manifestVars=manifests,
+                        latentVars=latents,
+                        mxPath(from=latents, to=manifests),
+                        mxPath(from=manifests, arrows=2),
+                        mxPath(from=latents, arrows=2, free=FALSE, values=1.0),
+                        mxData(observed=cov(demoOneFactor), type="cov", numObs=500)
+)
+
+factorFit1 <- mxRun(factorModel1)
+summary(factorFit1)
+```
+Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error as fingerprints of the entire objects will differ.
+For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
+Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1`. 
+Its `output` attribute contains timing variables `wallTime` and `cpuTime`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility:
+
+```{r}
+cat(factorFit1$output$wallTime)
+cat(factorFit1$output$cpuTime)
+```
+
+Thus, we recommend a checkpoint approach, in which reproducibility tests are used only in some meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size estimates, confidence intervals, Bayes factors, etc.).
+Specifically, we recommend to use at least the following checkpoints:
+- data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
+- preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
+- results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
+At the same time, avoid adding automated tests of entire statistical models but focus on the results.
+A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as the generic function `coef()` which can be used to extract point estimates from linear regression models or `parTable()` in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
+The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with OpenMx models), it may again contain timing information, which will not exactly reproduce.
+Last, note that some algorithm are not expected to reproduce identically but up to a certain precision.
+This is relevant for algorithm that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar.
+In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
+
 # Discussion

### 2024-06-25 manuscript.Rmd
added tidy perspective

@@
 In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
+Adepts of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `fit()` and `glance()` from the `broom` package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format, which lends itself for fingerprinting with our suggested approach.
 

### 2024-06-25 manuscript.Rmd
add broom ref

--- 2024-06-25 manuscript.Rmd
@@
 In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
+Adepts of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `fit()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format, which lends itself for fingerprinting with our suggested approach.
 

### 2024-06-25 manuscript.Rmd
include example of tidy/broom

--- 2024-06-25 manuscript.Rmd
@@
 In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
+Adepts of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
+Here is a brief example of how the `broom` package formats the output of a linear regression model:
+
+```{r}
+x <- rnorm(10)
+y <- rnorm(10)
+broom::tidy(lm(y~x))
+```
 

### 2024-06-25 manuscript.Rmd
revised abstract

@@
  However, a simple way to detect changes of these assets, i.e. irreproducibility, is still missing.
- We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically store meta data about original computational results and check later reproductions of the research assets for reproduction.
- With a minimal change to users' workflows, we hope that this approach increases transparency and trustworthiness of research assets.
+ We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically stores meta data about original computational results and verifies later reproduction attempts.
+ With a minimal change to users' workflows, we hope that this approach increases transparency and trustworthiness of digital research assets.
 

### 2024-06-25 manuscript.Rmd
fixed figure caption

@@
 
-```{r rstudioscreen1, fig.cap="Reproducibility report shown when a document is created for the first time and no reproducibility information is available yet."}
+```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight) instead of `R` as shown in this example R Markdown snippet.
+"}
 knitr::include_graphics("img/rstudio-screenshot-marker2.png")

### 2024-06-25 manuscript.Rmd
some more revision for clarity

@@
 Supported algorithm (through the digest package, @digest2022) are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
-To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers takes `r mean(benchmark$time)` seconds on our standard computers).
+To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers takes `r mean_time_s` seconds on our standard computers).
 We believe that in most cases, the speed factor will be irrelevant but it could become an issue when very large objects (e.g., entire neuroimaging data sets) are fingerprinted.
@@
 For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
-Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1`. 
-Its `output` attribute contains timing variables `wallTime` and `cpuTime`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility:
+Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables.
+Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility:
 
-```{r}
+```{r eval=TRUE, echo=TRUE}
 cat(factorFit1$output$wallTime)
@@
 
-Thus, we recommend a checkpoint approach, in which reproducibility tests are used only in some meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size estimates, confidence intervals, Bayes factors, etc.).
+Thus, we generally recommend a checkpoint approach, in which reproducibility tests are used only at selected, meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size point estimates, goodness-of-fit indices, standard error estimates, confidence intervals, test statistics, p values, Bayes factors, etc.).
 Specifically, we recommend to use at least the following checkpoints:
@@
 
-```{r}
+```{r echo=TRUE, eval=TRUE}
 x <- rnorm(10)

### 2024-06-26 manuscript.Rmd
added missing information fixed some refs

@@
 Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables.
-Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility:
+Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
+For illustration, here is the content of the `cpuTime` variable and the `timestump` variable for the aforementioned factor model, which was estimated when this manuscript was generated:
 
 ```{r eval=TRUE, echo=TRUE}
-cat(factorFit1$output$wallTime)
 cat(factorFit1$output$cpuTime)
+cat(factorFit1$output$timestamp)
 ```
@@
 In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
-Adepts of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
+Friends of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
 Here is a brief example of how the `broom` package formats the output of a linear regression model:
@@
 ```
+
+As a rule of thumb, we recommend to fingerprint the results of all parameter estimates from the `tidy()` function and all model fit statistics from the `glance()` function in a `reproducibleR` chunk.
 
@@
 
-- Some computations are executed and their results are stored in a JSON file as planned.
-Later on, someone changes the code in the R Markdown file, such that the results differ and a failure of reproduction is indicated.
+- Some computations are executed and their original results are stored in a JSON file as planned.
+Later on, someone modifies the R code in the R Markdown file, such that the results differ and a failure of reproduction is indicated.
+This change of code between the original computation and the reproduction attempt is caught by the package because fingerprints of the entire code chunk syntax are stored.
+Users are informed by a warning in the reproducibility report.
+The package will still try to reproduce each result and give individual reports on successes and failures.
 

### 2024-06-26 manuscript.Rmd
minor fixes

@@
  Various approaches have been suggested to increase chances of reproducibility.
- Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, presentation).
- However, a simple way to detect changes of these assets, i.e. irreproducibility, is still missing.
+ Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
+ However, a simple way to detect non-reproducibility, that is, unwanted changes in these assets over time is still missing.
  We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically stores meta data about original computational results and verifies later reproduction attempts.
@@
 Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
-A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a statistic or table in a Word document) [@peikert2021reproducible];
+A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) [@peikert2021reproducible];
 further, it is suitable to generate a variety of scientific assets other than manuscripts such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.

### 2024-06-26 manuscript.Rmd
some markdown fixes + one sentence

@@
 Then, all variables created in the process of a given data analysis will be tested for reproducibility.
-However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results will perfectly reproduce.
+However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results that researchers find meaningfull will perfectly reproduce.
 
@@
 ```
+
 Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error as fingerprints of the entire objects will differ.
@@
 Specifically, we recommend to use at least the following checkpoints:
+
 - data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
@@
 At the same time, avoid adding automated tests of entire statistical models but focus on the results.
+
 A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as the generic function `coef()` which can be used to extract point estimates from linear regression models or `parTable()` in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
@@
 Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
-
 - Original computations were executed and stored in the JSON data file.
@@
 To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
-
 - Some computations are executed and their original results are stored in a JSON file as planned.
@@
 The package will still try to reproduce each result and give individual reports on successes and failures.
-
 - `R` is not available anymore.
@@
 Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent.
-
 - A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken).

### 2024-06-26 manuscript.Rmd
added keywords

--- 2024-06-26 manuscript.Rmd
@@
 
-keywords          : "reproducibility, Markdown, R"
+keywords          : "reproducibility, R, Markdown, Open Science, statistical analysis, computation"
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')))`"

### 2024-06-26 manuscript.Rmd

--- 2024-06-26 manuscript.Rmd
@@
 Then, all variables created in the process of a given data analysis will be tested for reproducibility.
-However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results that researchers find meaningfull will perfectly reproduce.
+However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results that researchers find meaningful will perfectly reproduce.
 

### 2024-06-26 manuscript.Rmd
added Artner

@@
 Reproducibility is a prerequisite for replicable, robust and credible research in digital environments [@epskamp2019reproducibility; @hardwicke2018data].
+Surprisingly often, results from published raw data cannot be reproduced [@artner2021reproducibility].
  Various approaches have been suggested to increase chances of reproducibility of statistical data analyses in the R language [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac].
@@
 
-<!-- - The results of hash functions are generally not guaranteed to be stable over time. To trace back possible problems, -->
-
 ### Rigor in software development
@@
 It is not meant for storing large data and will blow up file size by a considerable factor (while at the same time losing numerical precision).
-At the same time, ensuring that the data set identical to the one used in the original analysis was used, adding a check on the contents of a `data.frame`.
+At the same time, ensuring that the data set is identical to the one used in the original analysis was used is essential (see our previous suggestion for checkpoints in a data analysis); that is, raw data in a `data.frame` or similar R structure should be tested for reproducibility but only indirectly using fingerprints.
+Note that many fitted statistical models in R also contain raw data, such as (general) linear model fits from the `stats` or `lme4` package or structural equation model fits from the `lavaan` or `OpenMx` packages.
+Repeating our previous suggestion, we recommend to only subject parameter estimates and fit indices to the reproducibility checks but not variables containing fitted statistical models.
 

### 2024-06-26 manuscript.Rmd
added further roles

@@
       - "Conceptualization"
+      - "Methodology"
+      - "Software"
       - "Writing - Original Draft Preparation"
@@
       - "Writing - Review & Editing"
+      - "Methodology"
 

### 2024-06-26 manuscript.Rmd
fixes in formatting options

@@
       background-color: #EEEEEE;'>
-      <h5>${title}</h5>${content}</div>"))
+      <h5>${title}</h5>
+      ${content}</div>"))
 ```
 
-For example, the following code reformats the appearance of the code report in PDF documents, such that the report is enclosed by two horizontal lines (\\hr elements), the title is displayed as a section header, there is a medium skip between title and content and content is displayed in a smaller font size:
+For example, the following code reformats the appearance of the code report in PDF documents (via \LaTeX), such that the report is enclosed by two horizontal lines (\\hrulefill elements), the title is displayed as a section header, there is a medium skip between title and content and content is displayed in a small font size:
 
@@
     options(reproducibleRchunks.templates = list(
-      pdf="\\hr \\section{${title}} 
-      \\medskip \\smaller 
-      ${content} \\hr"
+      latex="\\hrulefill \n \\section{${title}} 
+      \\medskip \\small 
+      ${content}\n \\hrulefill \n
+      ")
     )

### 2024-06-27 manuscript.Rmd
preprint version

@@
     institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research"
+
+note: Preprint submitted for publication
 
@@
 R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
-When ever the document is reproduced, all of its computer code chunks are executed and thus generate dynamic content in the document.
+Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
 This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated. 
-Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables and figures) in a specific journal style, such as APA style.
+Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables, figures, and statistical results) in a standardized style, such as APA style.
 Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
-A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors between results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) [@peikert2021reproducible];
-further, it is suitable to generate a variety of scientific assets other than manuscripts such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
+A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
+further, R Markdown is suitable to generate a variety of scientific assets other than manuscripts such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
 
-The `reproducibleRchunk` package allows users to make computational results in R Markdown documents automatically testable for reproduction (does the same script with the same data produce the same results, e.g.
-on a different computer and/or later in time) with minimal changes to your workflow.
-There are only two things users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks that should be automatically checked for reproducibility.
+The `reproducibleRchunk` package allows users to make computational results in R Markdown documents automatically testable for successful reproduction (does the same script with the same data produce the same results, e.g.
+on a different computer and/or later in time) with minimal changes to the workflow.
+There are only two things users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, they load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks for which reproduction should be automatically verified in future.
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a meta data file. 
@@
 Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
-The result of the reproduction attempt is displayed in a reproducibility report for each chunk.
-Customizations of these reports are possible as we will demonstrate later.
+Successes and failures of the reproduction attempts are displayed in a reproducibility report for each chunk.
+Customizations or suppression of these reports are possible as we will demonstrate later.
 A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
-Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over which stages of a data analysis or any other computation should be checked for reproducibility.
-For example, if there is non-reproducibility, is it due to changes in the data file, during data cleaning or statistical modeling. 
+Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over the stages of a data analysis or any other computation that should be checked for reproducibility.
+This allows for forensic examination of non-reproducibility; for example, finding out whether the non-reproducibility is due to changes in the data file, due to the data cleaning or the estimation of a statistical model.
 Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given statistical data analysis (or any other computation).
 Also, not all elements of a research asset may be required to reproduce (such as the current date displayed in a presentation of results).
-This enables more detailed forensics in case of irreproducibility.
-In the following, we will briefly demonstrate how successful and failed reproduction will be checked and displayed by the package, we will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
+In the following, we will briefly demonstrate how reproduction will be checked and how successful and failed attempts are displayed by the package.
+We will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
 
@@
 
-In this example, we assume that the R Markdown document is eventually knitted into a HTML-report (but all other pandoc-supported formats are supported as well).
+In this example, we assume that the R Markdown document is eventually knitted into a HTML-report (but all other pandoc-supported formats are supported as well, such as PDF or Word documents).
 On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file, which contains all information for further reproducibility tests. 
@@
 
-```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed."}
+```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed.", fig.pos="b"}
 knitr::include_graphics("img/generation-step2-failed.png")
@@
 
-In the following, we will describe how the package stores, retrieves and compares reproducibility information.
+In the following, we will describe how the package stores, retrieves and compares reproducibility information to verify reproduction attempts.
 First of all, the package executes `reproducibleR` code chunks just as regular R code chunks.
@@
 The name of the JSON file contains the original Markdown file and the chunk label and, by package default, starts with the prefix `.repro`. 
-That is, if reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored in a file called _.repro_sem_analysis.Rmd_datacleaning.json_.
-Once the document is re-generated and JSON data files exist, their content is checked against the newly computed chunk variables for identity.
+That is, reproducibility information of a chunk labelled `datacleaning` in the file `sem_analysis.Rmd` is stored in a file called _.repro_sem_analysis.Rmd_datacleaning.json_.
+Once the document is re-generated and matching JSON data files exist, their content is checked against the newly computed chunk variables for identity.
 
@@
 Fingerprints are realized via so-called one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
-Fingerprints allow for an efficient comparison of identity of results while avoid data security and privacy issues as they do not allow to reproduce the original data.
+Fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow to reproduce the original data.
 The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint (which is a 256 bit-long fingerprint that is usually displayed using 64 hexadecimal characters):
@@
 Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
-For illustration, here is the content of the `cpuTime` variable and the `timestump` variable for the aforementioned factor model, which was estimated when this manuscript was generated:
+For illustration, here is the content of the `cpuTime` variable and the `timestamp` variable for the aforementioned factor model, which was estimated when this manuscript was generated:
 
@@
 The output of the package is compatible with all pandoc output formats including the standard choices html, pdf, and docx. 
-Templates can be used to customize the output and summary functions allow users to generate arbitrary reproducibility reports either within the document or as separate files.
+Templates allow for customization of the appearance of reproducibility reports; further, the package allows users to generate entirely customized reproducibility reports either within a given R Mardkwon document or as separate files.
 
@@
 
-Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing reproducibility.
+Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing and verifying reproducibility.
 Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
 First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that execute code under identical conditions on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).
-On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., when Apple switched to their own processors) or because a user does not have admin privileges on a machine.
-Further some approaches rely on service providers such as the MRAN archive, which a while ago was unexpectedly terminated. 
-These different Achilles' heels now and then lead to situations, in which users will try to locally reproduce historic computational results and they may want to ensure that these reproductions were succesful.
-With our package, the success of such reproduction attempts becomes easily testable.
+On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., this happened when Apple switched to their own processor brand) or because a user does not have admin privileges for installing Docker in the first place.
+Further some approaches rely on service providers such as Microsoft's MRAN archive, which a while ago was unexpectedly terminated. 
+These different weak points now and then lead to situations, in which users will try to locally reproduce historic computational results and they may want to ensure that these reproductions were succesful.
+With our package, the success of such reproduction attempts becomes easily and formally testable.
 

### 2024-09-20 manuscript.Rmd
fixed corresponding address
added motivation

@@
 
-note: Preprint submitted for publication
+#note: Preprint submitted for publication
 
 authornote: |
-  Please address correspondence to: Prof. Dr. Andreas M. Brandmaier, Rüdesheimerstr. 50, 14197 Berlin, Germany. Email: andreas.brandmaier@medicalschool-berlin.de.
+  Please address correspondence to: Andreas M. Brandmaier, Lentzeallee 94, 14195 Berlin, Germany. Email: brandmaier@mpib-berlin.mpg.de.
   The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks.
@@
 
-The `reproducibleRchunk` package allows users to make computational results in R Markdown documents automatically testable for successful reproduction (does the same script with the same data produce the same results, e.g.
-on a different computer and/or later in time) with minimal changes to the workflow.
-There are only two things users need to change in their Markdown document if they are already using R Markdown: In the first code chunk of their document, they load the package and then change the code chunk type from `r` to `reproducibleR` for all code chunks for which reproduction should be automatically verified in future.
+When computations are repeated later in time with the goal of reproducing results computed earlier, it is hard and laborious to check whether results are identical.
+A typical case would be that readers download open data and open code accompanying a published report and re-run the code on their computer.
+In order to test whether these open materials reproduced, they would have to manually compare the newly computed results with the published manuscript -- and, often, it would not be immediately clear which results mapped to what parts of the manuscript.
+And even with dynamically-generated (R Markdown) documents, there is no straightforward way of checking reproducibility in an automated way.
+
+The `reproducibleRchunk` package answers this challenge and allows users to make computational results in R Markdown documents automatically testable for successful reproduction (does the same script with the same data produce the same results, e.g.
+on a different computer and/or later in time) with minimal changes to users' workflows.
+There are only two things users need to change in their Markdown document if they are already using R Markdown: (1) In the first code chunk of their document, they need to load the package, and (2) change the code chunk type from `r` to `reproducibleR` for every code chunk for which reproduction should be automatically verified in future.
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a meta data file. 

### 2025-02-14 manuscript.Rmd
some fixes based on reviews

@@
  Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
- However, a simple way to detect non-reproducibility, that is, unwanted changes in these assets over time is still missing.
- We introduce the R package `reproducibleRchunks`, which provides a new type of code chunk in R Markdown documents, which automatically stores meta data about original computational results and verifies later reproduction attempts.
+ However, a simple way to automatically verify reproducibility is still missing.
+ We introduce the R package `reproducibleRchunks`, which automatically stores metadata about original computational results and verifies later reproduction attempts automatically.
  With a minimal change to users' workflows, we hope that this approach increases transparency and trustworthiness of digital research assets.
@@
 library("reproducibleRchunks")
+library("knitr")
 r_refs("r-references.bib")
+#r_refs("r-references-anonymized.bib")
+```
+
+```{r}
+# Initialize an empty list to store legends
+legends <- list()
+
+# get hook
+original_plot_hook = knitr::knit_hooks$get('plot')
+
+# Define a hook to capture the figure legends
+knitr::knit_hooks$set(plot = function(x, options) {
+  if (!is.null(options$fig.cap)) {
+    legends <<- c(legends, options$fig.cap)
+  }
+  # Return the original plot hook
+  original_plot_hook(x, options)
+})
 ```
@@
  Many approaches rely on R Markdown [@rmarkdown2023].
-R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, python or others) in a single document [@xie2018r].
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, Python or others) in a single document [@xie2018r].
 Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
@@
 When computations are repeated later in time with the goal of reproducing results computed earlier, it is hard and laborious to check whether results are identical.
-A typical case would be that readers download open data and open code accompanying a published report and re-run the code on their computer.
+A typical case, in which verification of results is relevant, would be that readers or reviewers download open data and open code accompanying a manuscript and attempt to re-run the code on their computer.
 In order to test whether these open materials reproduced, they would have to manually compare the newly computed results with the published manuscript -- and, often, it would not be immediately clear which results mapped to what parts of the manuscript.
@@
 There are only two things users need to change in their Markdown document if they are already using R Markdown: (1) In the first code chunk of their document, they need to load the package, and (2) change the code chunk type from `r` to `reproducibleR` for every code chunk for which reproduction should be automatically verified in future.
-In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a meta data file. 
+In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a metadata file. 
 Furthermore, the code chunk itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
@@
 The core functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters).
-To use these new code chunks, the package must be loaded in the first regular `R` code chunk in the document using
+To use these new code chunks, the package must be loaded^[If an R Markdown file is knitted and the `reproducibleRchunks` package was not loaded, this error is shown: Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'`] in the first regular `R` code chunk in the document using
 
@@
 
-Once a document is (re-)generated with reproducibility data present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects amenable for reproducibility checks.
+Once a document is (re-)generated with metadata present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects amenable for reproducibility checks.
 Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
@@
 While virtually any R object is serializable in principle, we expect that the serialization algorithm is more likely to change over time for more complex objects in future than for simple objects.
-Using exact comparisons for testing compley object identity is sometimes more restrictive than one would expect.
+Using exact comparisons for testing complex objects' identity sometimes leads to unexpected results.
 This is particularly true for numeric comparisons. 
@@
 Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
-To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
+To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ten digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
 
@@
 
-In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents. First, to install the package, please install the latest version from our repository:
+In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents.
+First, to install the package, please install the latest version from CRAN^[The latest version can be found in the package repository: https://github.com/brandmaier/reproducibleRchunks]:
 
 ```{r eval=FALSE, echo=TRUE}
-if(!requireNamespace("pak"))install.packages("pak")
-pak::pkg_install("brandmaier/reproducibleRchunks")
+install.packages("reproducibleRchunks")
 ```
@@
 `reproducibleRchunks.digits`
-: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $8$.
+: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $10$.
 
@@
 
-### What could possible go wrong?
+### What could possibly go wrong?
 
@@
 
-- Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'` This error is shown if a R Markdown file is knitted and the `reproducibleRchunks` package was not loaded.
-Fix this by adding the statement `library(reproducibleRchunks)` in the first code chunk of the R Markdown document.
 - Original computations were executed and stored in the JSON data file.
@@
 The package will still try to reproduce each result and give individual reports on successes and failures.
+- The metadata files get lost.
+Without metadata files, no automated reproducibility check can be made.
+In this case, it may be a good idea to rerun the analysis in a computing environment as close as possible to the original computing environment and store the newly generated metadata files.
+It is advisable to manually compare the recreated results to some original reference (e.g., a report or published article) for consistency.
+- The `reproducibleRchunks` package is not available anymore.
+In this case, all R Markdown code chunks of type `reproducibleR` can be renamed to `R` and the manuscript will at least render even though without automated reproducibility checks.
+If there is a large number of code chunks, the following line of code can be inserted in the Markdown document in the very first code chunk, which will tell knitr to render all `reproducibleR` chunks as regular R chunks without the need to rename the chunks: `R knitr::knit_engines$set(reproducibleR = knitr::knit_engines$get("R")`.
 - `R` is not available anymore.
-As with every fashion, some day `R` will be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform.
+Some day `R` may be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform.
 Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent.
@@
 Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing and verifying reproducibility.
+For example, it would allow for reproducibility checks in peer review as suggested by @cruewell2023 could be automated using this package.
+According to the authors' convention, the current approach relies on testing whether results are _exactly reproducible_, which is consistent with the idea of awarding an Open Data badge.
+If there was a demand, the package could be extended to support the more fine-grained judgements suggested by @cruewell2023, such as _essentially reproducible_ (minor deviations in the decimals), _partially reproducible_ (minor deviations but the results were mostly numerically consistent),or _mostly not reproducible_ (major deviations).
+Note, however, that the consistent use of R Markdown already eliminates some sources of irreproducibility, such as copy-and-paste errors [@peikert2021reproducible].
 Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
@@
 :::
+
+# Contributions
+
+The authors made the following contributions.
+AB: Conceptualization, Methodology, Software, Writing - Original Draft Preparation, Writing - Review & Editing; AP: Writing - Review & Editing, Methodology.
+
+# Acknowledgements
+
+We thank Leonie Hagitte for providing comments on an earlier version of the manuscript.
+
+# Competing Interests
+
+The authors declare no competing interests.
+
+# Data accessibility statement
+
+The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks.
+The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.
+
+# Figure titles and legends
+
+```{r legends, results="asis"}
+for (i in 1:length(legends)) {
+  cat("Figure ",i ,".", legends[[i]],"\n\n")
+}
+
+```

### 2025-02-20 manuscript.Rmd
revision following reviewers' suggestions

@@
 Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments.
- Various approaches have been suggested to increase chances of reproducibility.
+ Various approaches have been suggested to increase chances of reproducibility of R code.
  Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
@@
 Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables, figures, and statistical results) in a standardized style, such as APA style.
-Also, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
+As one example, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
 A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
@@
 
-When computations are repeated later in time with the goal of reproducing results computed earlier, it is hard and laborious to check whether results are identical.
+When computations are repeated later in time with the goal of reproducing results computed earlier, it is hard and laborious to check whether results are identical or not. 
 A typical case, in which verification of results is relevant, would be that readers or reviewers download open data and open code accompanying a manuscript and attempt to re-run the code on their computer.
 In order to test whether these open materials reproduced, they would have to manually compare the newly computed results with the published manuscript -- and, often, it would not be immediately clear which results mapped to what parts of the manuscript.
-And even with dynamically-generated (R Markdown) documents, there is no straightforward way of checking reproducibility in an automated way.
+And even with dynamically-generated R Markdown documents, there is no straightforward way of checking reproducibility in an automated way.
 
-The `reproducibleRchunk` package answers this challenge and allows users to make computational results in R Markdown documents automatically testable for successful reproduction (does the same script with the same data produce the same results, e.g.
-on a different computer and/or later in time) with minimal changes to users' workflows.
-There are only two things users need to change in their Markdown document if they are already using R Markdown: (1) In the first code chunk of their document, they need to load the package, and (2) change the code chunk type from `r` to `reproducibleR` for every code chunk for which reproduction should be automatically verified in future.
+The ´reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
+This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
+This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
+We aim to first offer a less technical tutorial outlining a typical use case for the package. 
+Following this, we provide more detailed technical descriptions of the package's mechanics and various ways to customize its behavior.
+We believe that the package can be used effectively without a deep understanding of the technical details. 
+However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.
+
+For users already familiar with R Markdown, only two modifications are needed when using this package: 
+1. Load the reproducibleRchunk package in the first code chunk of your document
+2. Change the code chunk type from `r` to `reproducibleR` for each code chunk that requires automatic verification in the future.
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a metadata file. 
-Furthermore, the code chunk itself (that is, the code syntax) is also fingerprinted and stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
+Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
 Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
@@
 Customizations or suppression of these reports are possible as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+A straightforward way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown language for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over the stages of a data analysis or any other computation that should be checked for reproducibility.
@@
 
-The core functionality of the package is to provide a `reproducibleR` chunk, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters).
+The core functionality of the package is to provide  `reproducibleR` chunks, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters) but offer reproducibility checks in addition.
 To use these new code chunks, the package must be loaded^[If an R Markdown file is knitted and the `reproducibleRchunks` package was not loaded, this error is shown: Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'`] in the first regular `R` code chunk in the document using
@@
 
-In this example, we assume that the R Markdown document is eventually knitted into a HTML-report (but all other pandoc-supported formats are supported as well, such as PDF or Word documents).
-On document generation, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file, which contains all information for further reproducibility tests. 
-A reproducibility report is shown in the generated document below the code chunk indicating all variables of the given chunk whose data are stored.
-Figure \@ref(fig:rstudioscreen2) show a snippet from the HTML report generated on the first ever creation of the document.
+In this example, we assume that the R Markdown document is eventually converted into a HTML-report (but all other pandoc-supported formats are supported as well, such as PDF or Word documents).
+During initial document conversion, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file (with prefix _.repro_), which contains all information for further reproducibility verification tests.
+A reproducibility report is shown in the generated document below the code chunk indicating information about all variables of the given chunk.
+Figure \@ref(fig:rstudioscreen2) show a snippet from the HTML report generated during the initial creation of the document.
 
-Once the document is generated a second time and the reproducibility data exists, the computation is executed anew and its results are compared to the stored reproducibility information on a variable by variable basis.
-The reproducibility report will now contain information about the success of the individual reproduction attempts.
-In our example, the computation was reproduced successfully.
-A snippet from the resulting generated document is shown in Figure \@ref(fig:rstudioscreen3).
+Once the same document is reproduced, that is generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
+The reproducibility report will then detail the success of each reproduction attempt.
+In our example, the computation was successfully reproduced.
+A snippet from the resulting generated document is displayed in Figure \@ref(fig:rstudioscreen3).
 
@@
 
-If the document is regenerated and the computational result is not identical to the original result, this failure of reproduction is indicated in the reproducibility report.
+If the document is reproducedj and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
 In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
@@
 
+Ideally, the reproducibility metadata files are always kept together with the original Markdown file.
+For example, when providing Open Code on open platforms, the repository should contain both the Markdown files as well as the metadata files.
+Note that the metadata file format is particularly suitable to be used with version control systems [see @peikert2021reproducible].
+
 # Methods
 
-In the following, we will describe how the package stores, retrieves and compares reproducibility information to verify reproduction attempts.
-First of all, the package executes `reproducibleR` code chunks just as regular R code chunks.
-That is, there are no additional limits imposed on what can be computed and tested for reproducibility.
-Once code execution is completed, the package gathers information about all variables that were newly declared in the current chunk. 
+In the following, we describe how the package stores, retrieves and compares reproducibility information to verify reproduction attempts.
+First of all, the package executes `reproducibleR` code chunks just like regular R code chunks, meaning there are no additional restrictions on what can be computed and tested for reproducibility.
+After code execution, the package collects information about all variables that were newly declared in the current chunk. 
 The contents of those variables are stored in a separate JSON data file.
-JSON is a shorthand for Java Script Object Notation and its main features are that it is an open standard for mapping complex objects to text files with high simplicity and readability for machines and humans [@lennon2009introduction].
+JSON, short for JavaScript Object Notation, is an open standard for mapping complex objects to text files with high simplicity and readability for machines and humans [@lennon2009introduction].
 The name of the JSON file contains the original Markdown file and the chunk label and, by package default, starts with the prefix `.repro`. 
@@
 The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans. 
-Without going into detail, we believe it suffices to say that for each object, its type (character, integer, list, class, etc.), its value and its attributes are stored whereas values itself can again be collections of objects; in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
-For privacy reasons and to save disk space, we actually do not store the raw data by default but only fingerprints of the data.
-Fingerprints are realized via so-called one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
-Fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow to reproduce the original data.
-The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint (which is a 256 bit-long fingerprint that is usually displayed using 64 hexadecimal characters):
-
+Without going into detail, for each object, its type (e.g., character, integer, list, or class), its value and its attributes are stored.
+The stored values themselves values  can in turn consist of collections of objects (e.g., lists of lists); in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
+For privacy reasons and to save disk space, we actually do not store the raw data by default.
+Instead, we store "fingerprints" of the data.
+Digital fingerprints are fixed-length representations of an arbitrarily large digital object.
+They are generated such that small changes in the input object will lead to different fingerprints and, thus, they are ideally suited to verify integrity of the original data.
+The idea is very similar to a check sum digit in a credit card number, which can be computed from the remaining digits to check whether the credit card number is valid.
+Fingerprints are realized via one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
+As default, we chose the SHA256 algorithm, which allows us to map any content of a variable in R to a 256 bit-long fingerprint or "hash", which is usually displayed using 64 hexadecimal characters.
+The following shows the SHA256-fingerprint of the content of variable `numbers` from the previous example:
 \begin{verbatim}
@@
 \end{verbatim}
+And the following example is a SHA256 fingerprint of the first chapter of Genesis from the Old Testament:
+\begin{verbatim}
+```{r echo=FALSE, eval=TRUE, results="asis"}
+library(logos)
+genesis <- select_passage("Genesis",chapter = 1, language = "English", testament = "old")
+content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
+content <- gsub(pattern="\\n",replacement="  \n", x=content)
+cat(content)
+```
+\end{verbatim}
+Note that these exemplary fingerprints have the same length despite both variables having different variable types and sizes.
+Due to their small size, fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow the original objects to be reproduced.
 
@@
 
-Once a document is (re-)generated with metadata present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects amenable for reproducibility checks.
-Identity of original and reproduced results are checked with `identical()` function from R's `base` package, which according to the R documentation is a "safe and reliable way to test two objects (even complex ones) for being exactly identical" (see `?identical` R documentation).
+Once a document is (re-)generated with metadata present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects that are amenable for reproducibility checks.
+Identicality of original and reproduced results is checked with the `all.equal()` function from R's `base` package, which according to the R documentation is a "is a utility to compare R objects x and y testing ‘near equality’.
+If they are different, comparison is still made to some extent, and a report of the differences is returned." (see `?all.equal` R documentation).
 
@@
 
-What types of objects are suited for automated reproducibility tests? The answer is simple. In principle, any R object is suitable no matter how complex.
+In principle, any R object is suitable for reproducibility testing no matter how complex.
 For creating fingerprints, the package uses R's serialization methods, which take any R object as input.
@@
 ```
-
-Yet, we recommend subjecting simple objects to reproducibility checks, that is, preferably numeric variables such as single numbers or vectors of numbers. 
-In a typical statistical data analysis this would correspond to, for example, parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
-While virtually any R object is serializable in principle, we expect that the serialization algorithm is more likely to change over time for more complex objects in future than for simple objects.
-Using exact comparisons for testing complex objects' identity sometimes leads to unexpected results.
-This is particularly true for numeric comparisons. 
-For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
-Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
-To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ten digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
 
@@
 
-In practice, we expect cases, in which the reproduction report should be displayed (e.g., when developing a data analysis script or reproducing a former one) or not (e.g., when rendering a presentation or a manuscript for submission to a journal).
+In practice, we expect that reproducible R chunks will be used both in cases, in which the reproduction report should be displayed (e.g., when developing a data analysis script or reproducing a former one) and those where it should not (e.g., when rendering a presentation or a manuscript for submission to a journal).
 By default, reproduction report statements are produced for each chunk but this default can be changed via the code chunk argument `report=FALSE`.
@@
 This information is returned by function `get_reproducibility_summary()` which returns a `data.frame` with three columns. 
-The first column contains the name of the code chunk, the second column contains a variable name and the third column contains a boolean variable representing whether the reproduction was a success. 
+The first column contains the name of the code chunk, the second column contains a variable name, and the third column contains a boolean variable representing whether the reproduction was a success. 
 This information can be used to either generate custom reports, such as one general report at the very end of the document, or write reports to a separate file.
@@
 
-For example, the following code reformats the appearance of the code report in PDF documents (via \LaTeX), such that the report is enclosed by two horizontal lines (\\hrulefill elements), the title is displayed as a section header, there is a medium skip between title and content and content is displayed in a small font size:
+For example, the following code reformats the appearance of the code report in PDF documents (via LaTeX), such that the report is enclosed by two horizontal lines (\\hrulefill elements), the title is displayed as a section header, there is a medium skip between title and content, and content is displayed in a small font size:
 
@@
 In the following, we give recommendations on how to use the package in typical cases of statistical data analysis and reporting.
-It is possible to simply replace all classic `R` code chunks with `reproducibleR` code chunks.
+A simple way to start is to replace all classic `R` code chunks with `reproducibleR` code chunks.
 Then, all variables created in the process of a given data analysis will be tested for reproducibility.
@@
 
-A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as the generic function `coef()` which can be used to extract point estimates from linear regression models or `parTable()` in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
+A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
+To this end, several packages support the generic function `coef()`, which can be used, for example, to extract point estimates from linear regression models.
+Other packages offer their own accessor function such as `parTable()` for structural equation models in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
 The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with OpenMx models), it may again contain timing information, which will not exactly reproduce.
-Last, note that some algorithm are not expected to reproduce identically but up to a certain precision.
-This is relevant for algorithm that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar.
-In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
-Friends of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
+Last, note that sometimes there may small deviations of exact reproducibility, which can be tolerated.
+For example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit), numeric representations may count as reproducible if they are identical up to some numerical precision.
+To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ten digits).
+Lastly, users may wish to test identity of reproduction up to a certain precision less than the package default.
+This could be relevant for algorithms that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar (even though, their perfect reproducibility should usually be guarantueed by setting a random seed).
+In this case, approximate reproducibility tests could be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
+Adherents of the tidyverse approach [@wickham2019welcome] are well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package.
+These functions convert parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easily accessible format known as _tibble_.
+This format is particularly suitable for fingerprinting using our proposed approach.
 Here is a brief example of how the `broom` package formats the output of a linear regression model:
@@
 The output of the package is compatible with all pandoc output formats including the standard choices html, pdf, and docx. 
-Templates allow for customization of the appearance of reproducibility reports; further, the package allows users to generate entirely customized reproducibility reports either within a given R Mardkwon document or as separate files.
+Templates allow for customization of the appearance of reproducibility reports; further, the package allows users to generate entirely customized reproducibility reports either within a given R Markdown document or as separate files.
 
-### What could possibly go wrong?
+### Potential long-term issues
 
-In the following, we briefly discuss a few situations, in which threats to reproducibility happened and explain how the package handles these situations by default:
+In the following, we briefly discuss a few future situations, in which threats to reproducibility or reproducibility checking happened and explain how these situations can be dealt with:
 
@@
 - A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken).
-The version of the serialize protocol can be manually overwritten to an older version if newer versions should not be backward compatible.
+The metadata contains information about the R version used to generate the metadata, so future versions of our package could easily be adapted to this change.
 As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2024) using this command: `options(serializeVersion=2)`.
@@
 On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., this happened when Apple switched to their own processor brand) or because a user does not have admin privileges for installing Docker in the first place.
-Further some approaches rely on service providers such as Microsoft's MRAN archive, which a while ago was unexpectedly terminated. 
+Further, some approaches rely on service providers such as Microsoft's MRAN archive, which a while ago was unexpectedly terminated. 
 These different weak points now and then lead to situations, in which users will try to locally reproduce historic computational results and they may want to ensure that these reproductions were succesful.

### 2025-02-20 manuscript.Rmd
some more changes during revision

@@
 We aim to first offer a less technical tutorial outlining a typical use case for the package. 
-Following this, we provide more detailed technical descriptions of the package's mechanics and various ways to customize its behavior.
+Following this, we provide more detailed technical background information about the package and various ways to customize its behavior.
 We believe that the package can be used effectively without a deep understanding of the technical details. 
@@
 Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_.
+The entire tutorial file can be found in the R package in sub directory "inst/examples_simple_demo" or can be downloaded from \url{https://github.com/brandmaier/reproducibleRchunks/tree/main/inst/examples/simple_demo}.
 In the chunk, a new variable `my_sum` is declared and defined to be the sum of some variable `x` plus one. 
@@
 
-If the document is reproducedj and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
+If the document is reproduced and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
 In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
@@
 Note that the metadata file format is particularly suitable to be used with version control systems [see @peikert2021reproducible].
+If users intentionally modify R code later, e.g., because they want to fix an error, then the metadata files would need to be deleted and recreated.
+Version control can help users in documenting such changes.
 
@@
 ```{r echo=FALSE, eval=TRUE, results="asis"}
-content <- jsonlite::serializeJSON(digest::digest(numbers,algo = "sha256"),pretty=TRUE)
+#content <- jsonlite::serializeJSON(digest::digest(numbers,algo = "sha256"),pretty=TRUE)
+content <- digest::digest(numbers,algo = "sha256")
 content <- gsub(pattern="\\n",replacement="  \n", x=content)
@@
 genesis <- select_passage("Genesis",chapter = 1, language = "English", testament = "old")
-content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
+#content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
+content <- digest::digest(genesis,algo = "sha256")
 content <- gsub(pattern="\\n",replacement="  \n", x=content)

### 2025-03-01 manuscript.Rmd
add my comments

@@
 
-The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans. 
-Without going into detail, for each object, its type (e.g., character, integer, list, or class), its value and its attributes are stored.
-The stored values themselves values  can in turn consist of collections of objects (e.g., lists of lists); in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
-For privacy reasons and to save disk space, we actually do not store the raw data by default.
-Instead, we store "fingerprints" of the data.
-Digital fingerprints are fixed-length representations of an arbitrarily large digital object.
-They are generated such that small changes in the input object will lead to different fingerprints and, thus, they are ideally suited to verify integrity of the original data.
-The idea is very similar to a check sum digit in a credit card number, which can be computed from the remaining digits to check whether the credit card number is valid.
-Fingerprints are realized via one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
-As default, we chose the SHA256 algorithm, which allows us to map any content of a variable in R to a 256 bit-long fingerprint or "hash", which is usually displayed using 64 hexadecimal characters.
-The following shows the SHA256-fingerprint of the content of variable `numbers` from the previous example:
+The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans, which can ease "forensics" in case reproduction fails.
+Such forensic investigation is further facilitated, the package also stores the code syntax.
+While JSON can represent arbitrarily complex R objects (such as an entire regression model), we opted to not store all computational results by default but only their "fingerprints".
+Using the fingerprint, we can determine if an object has changed without storing the object itself.
+This strategy avoids accidental leakage of potentially sensitive data and reduces storage demands.
+Fingerprints are realized via so-called one-way hash functions (we use the SHA256 algorythm by default), which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
+<!---AP: I think this section could be removed: start--->
+Fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow to reproduce the original data.
+The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint (which is a 256 bit-long fingerprint that is usually displayed using 64 hexadecimal characters):
+
 \begin{verbatim}
@@
 This allows later checks whether a potential non-reproduction is due to a modification of the code chunk since the original computation.
+
+<!---AP: I think this section could be removed: end--->
+
 The storage of the fingerprinted information about a code chunk and its computational results is illustrated in the schematic in Figure \@ref(fig:jsonschema).
@@
 ```
+
+<!---I think this is sound advice and should be kept--->
+Yet, we recommend subjecting simple objects to reproducibility checks, that is, preferably numeric variables such as single numbers or vectors of numbers. 
+In a typical statistical data analysis this would correspond to, for example, parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
+Using exact comparisons for testing complex object identity is sometimes more restrictive than one would expect.
+Complex objects (like the fit objects of statistical models, or plots), often contain information that changes on each run, like time stamps, surprising randomness, numerical instability etc.
+which makes the object technically not identical but would rarely count as irreproducible.
+Another complication is numeric comparisons. 
+For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
+Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
+To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
 

### 2025-03-03 manuscript.Rmd
revision contd

@@
 authornote: |
-  Please address correspondence to: Andreas M. Brandmaier, Lentzeallee 94, 14195 Berlin, Germany. Email: brandmaier@mpib-berlin.mpg.de.
+  Please address correspondence to: Andreas M. Brandmaier, Rüdesheimer Str. 50, 14197 Berlin, Germany. Email: andreas.brandmaier@medicalschool-berlin.de.
   The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks.
@@
  Many approaches rely on R Markdown [@rmarkdown2023].
-R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, what are list elements) and computer code (R, Python or others) in a single document [@xie2018r].
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements) and computer code (R, Python, or others) in a single document [@xie2018r].
 Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
 This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated.
-Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables, figures, and statistical results) in a standardized style, such as APA style.
-As one example, (R) Markdown forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
+Packages like `papaja` [@R-papaja] or `stargazer` [@R-stargazer] further extend R Markdown's capabilities to format documents (including tables, figures, and statistical results) in a standardized style, such as APA style [@apa].
+As another example, (R) Markdown also forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
 A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
-further, R Markdown is suitable to generate a variety of scientific assets other than manuscripts such as posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible] from the same Markdown language.
+further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible], all using the same Markdown approach.
 
@@
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a metadata file. 
-Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computation.
+Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computational results.
 Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
@@
 A straightforward way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown language for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
-Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over the stages of a data analysis or any other computation that should be checked for reproducibility.
+Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at what stages during the data analysis.
 This allows for forensic examination of non-reproducibility; for example, finding out whether the non-reproducibility is due to changes in the data file, due to the data cleaning or the estimation of a statistical model.
@@
 If the document is reproduced and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
-In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
+In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
 
@@
 
-This vector can be serialized in raw format to a JSON format as follows:
+This vector is serialized in raw format to a JSON format as follows:
 
@@
 
-The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans, which can ease "forensics" in case reproduction fails.
-Such forensic investigation is further facilitated, the package also stores the code syntax.
+The JSON format has a clear, formal structure that facilitates parsing of information for both computers and humans.
+This accessibility can ease forensics if reproduction issues arise.
+To further aid forensic investigation, the package also stores information about the code syntax used to generate an object.
 While JSON can represent arbitrarily complex R objects (such as an entire regression model), we opted to not store all computational results by default but only their "fingerprints".
@@
 Fingerprints are realized via so-called one-way hash functions (we use the SHA256 algorythm by default), which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
-<!---AP: I think this section could be removed: start--->
-Fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow to reproduce the original data.
 The following shows how the information about variable `numbers` is stored using a SHA256-fingerprint (which is a 256 bit-long fingerprint that is usually displayed using 64 hexadecimal characters):
-
 \begin{verbatim}
@@
 \end{verbatim}
-And the following example is a SHA256 fingerprint of the first chapter of Genesis from the Old Testament:
-\begin{verbatim}
-```{r echo=FALSE, eval=TRUE, results="asis"}
-library(logos)
-genesis <- select_passage("Genesis",chapter = 1, language = "English", testament = "old")
-#content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
-content <- digest::digest(genesis,algo = "sha256")
-content <- gsub(pattern="\\n",replacement="  \n", x=content)
-cat(content)
-```
-\end{verbatim}
-Note that these exemplary fingerprints have the same length despite both variables having different variable types and sizes.
-Due to their small size, fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow the original objects to be reproduced.
-
-In addition to the contents of all newly declared variables in a `reproducibleR`-chunk, the package also stores a fingerprint of the code syntax.
-This allows later checks whether a potential non-reproduction is due to a modification of the code chunk since the original computation.
-
-<!---AP: I think this section could be removed: end--->
-
 The storage of the fingerprinted information about a code chunk and its computational results is illustrated in the schematic in Figure \@ref(fig:jsonschema).
@@
 In principle, any R object is suitable for reproducibility testing no matter how complex.
-For creating fingerprints, the package uses R's serialization methods, which take any R object as input.
+For creating fingerprints, the package uses R's serialization methods, which take any R object as input and prepares it for storage.
 That is, all of the following variables (`x` of class _integer_, `y` of class _character_, `z` of class _lm_, and `lst` of class _list_) are valid examples for the automated reproducibility checks:
@@
 ```
-
-<!---I think this is sound advice and should be kept--->
-Yet, we recommend subjecting simple objects to reproducibility checks, that is, preferably numeric variables such as single numbers or vectors of numbers. 
-In a typical statistical data analysis this would correspond to, for example, parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
-Using exact comparisons for testing complex object identity is sometimes more restrictive than one would expect.
-Complex objects (like the fit objects of statistical models, or plots), often contain information that changes on each run, like time stamps, surprising randomness, numerical instability etc.
-which makes the object technically not identical but would rarely count as irreproducible.
-Another complication is numeric comparisons. 
-For example results of numeric computations may be logically equivalent but not bitwise equivalent; in numeric representations, there is sometimes a bit reserved for the sign of a number, thus it can be possible to have -0 and +0, which both are exactly zero in a decimal representation but the binary representations differ.
-Numeric representations may also differ up to a given numeric precision, for example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit).
-To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, eight digits) but this mechanism only works for plain numbers/matrices and possible not for complex composite objects.
 
@@
 If the R markdown document is rendered a second time, the computational results are recomputed and compared against the original results. 
-For each result, success or failure or reproduction is reported separately.
+For each result, success or failure of reproduction is reported separately.
 
@@
 `reproducibleRchunks.hashing_algorithm`
-: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithm offered by the `digest` package are used.
+: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithms offered by the `digest` package are used.
 
@@
 First, it is possible to change the precision with which numeric results are stored.
-By default, this is up to 8 digits.
+By default, this is up to 10 digits.
 The following line reduces the precision to only four digits after the decimal point:
@@
 In this application, the choice of algorithm is not overly crucial.
-Supported algorithm (through the digest package, @digest2022) are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
+Supported algorithm [provided by the digest package, @digest2022] are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
 To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers takes `r mean_time_s` seconds on our standard computers).
@@
 Then, all variables created in the process of a given data analysis will be tested for reproducibility.
-However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all results that researchers find meaningful will perfectly reproduce.
+However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all statistical results that researchers find meaningful will perfectly reproduce.
 
@@
 
-Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error as fingerprints of the entire objects will differ.
+Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error.
 For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
 Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables.
-Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
+Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
 For illustration, here is the content of the `cpuTime` variable and the `timestamp` variable for the aforementioned factor model, which was estimated when this manuscript was generated:

### 2025-03-04 manuscript.Rmd
a whole lot of love for the references

@@
 
-The ´reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
+The `reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
 This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
@@
 
-```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight) instead of `R` as shown in this example R Markdown snippet.
-"}
+```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight)"}
 knitr::include_graphics("img/rstudio-screenshot-marker2.png")
@@
 
-```{r rstudioscreen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is generated for the first time."}
+```{r rstudioscreen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is generated for the first time."}
 knitr::include_graphics("img/generation-step1.png")
@@
 
-```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and a variable is successfully reproduced."}
+```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is re-generated and a result is successfully reproduced."}
 knitr::include_graphics("img/generation-step2.png")
@@
 
-```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed.", fig.pos="b"}
+```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a result failed.", fig.pos="b"}
 knitr::include_graphics("img/generation-step2-failed.png")
@@
 The metadata contains information about the R version used to generate the metadata, so future versions of our package could easily be adapted to this change.
-As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2024) using this command: `options(serializeVersion=2)`.
+As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2025) using this command: `options(serializeVersion=2)`.
 

### 2025-04-01 manuscript.Rmd
changes in response to reviews

@@
  Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
- However, a simple way to automatically verify reproducibility is still missing.
- We introduce the R package `reproducibleRchunks`, which automatically stores metadata about original computational results and verifies later reproduction attempts automatically.
+However, a simple way to automatically verify reproducibility of such R Markdown documents is still missing.
+ We introduce the R package reproducibleRchunks, which automatically stores metadata about original computational results in an R Markdown document and verifies later reproduction attempts automatically.
  With a minimal change to users' workflows, we hope that this approach increases transparency and trustworthiness of digital research assets.
@@
 
+# Package at a glance
 For users already familiar with R Markdown, only two modifications are needed when using this package: 
@@
 Using the fingerprint, we can determine if an object has changed without storing the object itself.
-This strategy avoids accidental leakage of potentially sensitive data and reduces storage demands.
+This strategy avoids accidental leakage of potentially sensitive data^[Note that many fitted statistical models in R also contain raw data, such as (general) linear model fits from the `stats` or `lme4` package or structural equation model fits from the `lavaan` or `OpenMx` packages.] and reduces storage demands.
 Fingerprints are realized via so-called one-way hash functions (we use the SHA256 algorythm by default), which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
@@
 
-We introduced an R package that allows for R code chunks in R Markdown documents that are automatically tested for reproducibility. 
-The output of the package is compatible with all pandoc output formats including the standard choices html, pdf, and docx. 
+## Summary
+
+Dynamic document generation using R Markdown (e.g., reports, posters, or presentations) is an important building block for reproducible research. 
+However, a simple way to automatically verify reproducibility of dynamically generated content within such R Markdown documents is still missing. 
+We developed an R package that enables automatic reproducibility testing of R code chunks within R Markdown documents while requiring essentially no changes to users' workflows.
+This is achieved by storing meta-data about computational results, which can later be used to verify reproducibility.
 Templates allow for customization of the appearance of reproducibility reports; further, the package allows users to generate entirely customized reproducibility reports either within a given R Markdown document or as separate files.
 
-### Potential long-term issues
+## Strength and practical implications
+
+Adopting reproducibleR chunks in a workflow offers several benefits.
+First of all, it allows to verify reproducibility manually whenever an R Markdown document is rendered. 
+This may be particularly relevant when collaboratively working on scientific reports. 
+Adopting the proposed approach ensures that the same results are generated across all computers of a team of researchers. 
+Even in projects where a person works alone, the proposed workflow is useful. 
+It ensures that all computed results remain consistent throughout the lifecycle of a scientific report, from the first draft through peer-review phases to publication.
+In addition to manually verifying computational results, automatic checks are also possible.
+Documents can be systematically checked for reproducibility using the function call `isReproducible()`, which takes as input an R Markdown file and returns a logical value indicating whether reproduction was successful.
+On collaborative platforms like GitHub, reproducibility could be automatically tested with a GitHub action whenever the repository is modified. 
+It's conceivable that journals could use a similar mechanism to check the reproducibility of a paper upon submission.
+
+## Limitations
+
+Our approach is currently limited to users who rely on R Markdown. 
+Regular R scripts are not amenable to our proposed approach.
+Recently, Quarto [@quarto2024] was proposed as a next-generation Markdown publishing system and it has been gaining popularity. 
+However, our proposed approach is not yet compatible with Quarto.
+In general, our package relies on the availability and stability of some R functions and packages used for serializing objects and generating fingerprints.
+Below, we outline these issues and other short-term and long-term issues one may experience with this package.
+
+### Potential short-term and long-term issues
 
@@
 - Original computations were executed and stored in the JSON data file.
-Later, a reproduction attempt is made using the exact same R Markdown file but on a different computer.
+Later, a reproduction attempt fails using the exact same R Markdown file executed on a different computer.
 This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021reproducible].
@@
 - The `reproducibleRchunks` package is not available anymore.
-In this case, all R Markdown code chunks of type `reproducibleR` can be renamed to `R` and the manuscript will at least render even though without automated reproducibility checks.
+In this case, all R Markdown code chunks of type `reproducibleR` can be renamed to `R`, allowing the manuscript to render, albeit without automated reproducibility checks.
 If there is a large number of code chunks, the following line of code can be inserted in the Markdown document in the very first code chunk, which will tell knitr to render all `reproducibleR` chunks as regular R chunks without the need to rename the chunks: `R knitr::knit_engines$set(reproducibleR = knitr::knit_engines$get("R")`.
-- `R` is not available anymore.
-Some day `R` may be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform.
-Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent.
 - A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken).
@@
 
-### Rigor in software development
+## Rigor in software development
 
@@
 
-### Storing and publishing reproducibility information
-
-We strongly discourage making reproducibility checks on complete raw data. 
-This is mainly for two reasons.
-First, there may arise unwanted data protection issues when raw data is (accidentally) stored, copied or published via the reproducibility JSON files.
-Second, the JSON format is inefficient in terms of storage space required.
-It is designed to be human-readable and conformable with versioning tools such as git.
-It is not meant for storing large data and will blow up file size by a considerable factor (while at the same time losing numerical precision).
-At the same time, ensuring that the data set is identical to the one used in the original analysis was used is essential (see our previous suggestion for checkpoints in a data analysis); that is, raw data in a `data.frame` or similar R structure should be tested for reproducibility but only indirectly using fingerprints.
-Note that many fitted statistical models in R also contain raw data, such as (general) linear model fits from the `stats` or `lme4` package or structural equation model fits from the `lavaan` or `OpenMx` packages.
-Repeating our previous suggestion, we recommend to only subject parameter estimates and fit indices to the reproducibility checks but not variables containing fitted statistical models.
-
-### Summary
+## Outlook
 

### 2025-04-01 manuscript.Rmd
more revision based on reviews

@@
 output            : papaja::apa6_pdf
+#output            : papaja::apa6_docx
 header-includes:
@@
 Reproducibility is a prerequisite for replicable, robust and credible research in digital environments [@epskamp2019reproducibility; @hardwicke2018data].
-Surprisingly often, results from published raw data cannot be reproduced [@artner2021reproducibility].
- Various approaches have been suggested to increase chances of reproducibility of statistical data analyses in the R language [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac].
+Surprisingly often, results from published raw data cannot be reproduced.
+For example, @artner2021reproducibility, were able to reproduce only 70% of major scientific conclusions from a sample of articles published in Psychology -- however, the problem is likely pervasive in all scientific disciplines.
+ Various approaches have been suggested to increase chances of reproducibility of statistical data analyses in the R language [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac].
  Many approaches rely on R Markdown [@rmarkdown2023].
-R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements) and computer code (R, Python, or others) in a single document [@xie2018r].
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements) and executable computer code (R, Python, or other languages) in a single document [@xie2018r].
 Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
@@
 A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
-further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021reproducible], all using the same Markdown approach.
+further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021samething], all using the same Markdown approach.
 
-When computations are repeated later in time with the goal of reproducing results computed earlier, it is hard and laborious to check whether results are identical or not. 
-A typical case, in which verification of results is relevant, would be that readers or reviewers download open data and open code accompanying a manuscript and attempt to re-run the code on their computer.
-In order to test whether these open materials reproduced, they would have to manually compare the newly computed results with the published manuscript -- and, often, it would not be immediately clear which results mapped to what parts of the manuscript.
-And even with dynamically-generated R Markdown documents, there is no straightforward way of checking reproducibility in an automated way.
+In sum, using R Markdown for the creation of scientific reports including statistical computations comes with various advantages, particularly avoiding copy & paste errors as a threat to reproducibility.
+However, there are still various other sources of errors that are not eliminated solely by using R Markdown.
+For example, different R versions may lead to different results; a famous example is the change in the random number generator between R versions [@peikert2021reproducible].
+For another example, bugfixes or changes in package defaults across different package versions may go unnoticed and lead to different results.
+Therefore, it is essential to monitor changes in computational results within R Markdown documents throughout the document's entire lifecycle -- from the initial draft, through (peer-)review, to publication and long-term storage.
+Currently, there is no straightforward, automated method for documenting and verifying reproducibility.
+The `reproducibleRchunks` package addresses this challenge by ensuring that computational results in R Markdown documents are automatically testable for successful reproduction. 
+It verifies whether the same script with the same data produces identical results, even on different computers or at later times, with minimal disruption to users' workflows.
 
-The `reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
-This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
-This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
-We aim to first offer a less technical tutorial outlining a typical use case for the package. 
-Following this, we provide more detailed technical background information about the package and various ways to customize its behavior.
-We believe that the package can be used effectively without a deep understanding of the technical details. 
-However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.
+This article is written for readers who use R for statistical analyses and ideally have some experience with R Markdown. 
+We begin with a straightforward tutorial demonstrating a typical use case for the package. 
+Following this, we delve into more detailed technical background information and exokire various ways to customize the package's behavior.
+While a deep understanding of the technical details isn't necessary for effective use, gaining insight into these aspects can enhance the user experience and help readers appreciate the challenges of ensuring reproducibility.
 
 # Package at a glance
-For users already familiar with R Markdown, only two modifications are needed when using this package: 
-1. Load the reproducibleRchunk package in the first code chunk of your document
-2. Change the code chunk type from `r` to `reproducibleR` for each code chunk that requires automatic verification in the future.
+For users already familiar with R Markdown, only two modifications are needed when using the package:
+
+1. Load the `reproducibleRchunk` package in the first code chunk of your document
+2. Change the code chunk type from `r` to `reproducibleR` for every code chunk whose reproducibility should be automatically verified.
+
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a metadata file. 
@@
 Customizations or suppression of these reports are possible as we will demonstrate later.
-A straightforward way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown language for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
-Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at what stages during the data analysis.
+A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown language for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at which stages during the data analysis.
 This allows for forensic examination of non-reproducibility; for example, finding out whether the non-reproducibility is due to changes in the data file, due to the data cleaning or the estimation of a statistical model.
@@
 
-Once the same document is reproduced, that is generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
+Once the document is reproduced, that is, generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
 The reproducibility report will then detail the success of each reproduction attempt.
@@
 set.seed(345)
-numbers_in_simulation <- 1000
+numbers_in_simulation <- 10000
 benchmark <- microbenchmark::microbenchmark(digest::sha1(numbers), unit = "seconds", check = "equal", setup = {numbers <- runif(numbers_in_simulation)})
@@
 
-- data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
-- preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
-- results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
+- Data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
+- Preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
+- Results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
 At the same time, avoid adding automated tests of entire statistical models but focus on the results.
@@
 Later, a reproduction attempt fails using the exact same R Markdown file executed on a different computer.
-This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021reproducible].
+This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021samething].
 Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility.
-To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
+To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
 - Some computations are executed and their original results are stored in a JSON file as planned.
@@
 If there was a demand, the package could be extended to support the more fine-grained judgements suggested by @cruewell2023, such as _essentially reproducible_ (minor deviations in the decimals), _partially reproducible_ (minor deviations but the results were mostly numerically consistent),or _mostly not reproducible_ (major deviations).
-Note, however, that the consistent use of R Markdown already eliminates some sources of irreproducibility, such as copy-and-paste errors [@peikert2021reproducible].
-Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
+Note, however, that the consistent use of R Markdown already eliminates some sources of irreproducibility, such as copy-and-paste errors [@peikert2021samething].
+Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
 First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that execute code under identical conditions on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).

### 2025-04-03 manuscript.Rmd

@@
  Various approaches have been suggested to increase chances of reproducibility of R code.
- Many of them rely on R Markdown as a language to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
+ Many of them rely on R Markdown as a tool to dynamically generate reproducible research assets (e.g., reports, posters, or presentations).
 However, a simple way to automatically verify reproducibility of such R Markdown documents is still missing.
@@
 In `reproducibleR` chunks, all newly declared variables are automatically identified and their contents are stored in a metadata file. 
-Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computational results.
+Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computational environment.
 Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
@@
 Customizations or suppression of these reports are possible as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown language for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at which stages during the data analysis.
@@
 
-The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks.
+The R package described in this article can be downloaded from https://github.com/brandmaier/reproducibleRchunks or https://cran.r-project.org/web/packages/reproducibleRchunks/.
 The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.

### 2025-04-03 manuscript.Rmd
first round of edits suggested by JD

@@
 The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.
-We thank Leonie Hagitte for providing comments on an earlier version of the manuscript.
+We thank Julia Delius forher helpful assistance in language and style editing.
+We also thank Leonie Hagitte for providing comments on an earlier version of the manuscript.
 
@@
  Computational results are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
-Reproducibility is a prerequisite for replicable, robust and transparent research in digital environments.
+Reproducibility is a prerequisite for replicable, robust, and transparent research in digital environments.
  Various approaches have been suggested to increase chances of reproducibility of R code.
@@
 Computational results, including those from statistical data analyses, are considered _reproducible_ if the same computation on the same data yields the same results if performed on a different computer or on the same computer later in time.
-Reproducibility is a prerequisite for replicable, robust and credible research in digital environments [@epskamp2019reproducibility; @hardwicke2018data].
+Reproducibility is a prerequisite for replicable, robust, and credible research in digital environments [@epskamp2019reproducibility; @hardwicke2018data].
 Surprisingly often, results from published raw data cannot be reproduced.
-For example, @artner2021reproducibility, were able to reproduce only 70% of major scientific conclusions from a sample of articles published in Psychology -- however, the problem is likely pervasive in all scientific disciplines.
+For example, @artner2021reproducibility, were able to reproduce only 70% of major scientific conclusions from a sample of articles published in the domain of Psychology -- however, the problem is likely pervasive in all scientific disciplines.
  Various approaches have been suggested to increase chances of reproducibility of statistical data analyses in the R language [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac].
  Many approaches rely on R Markdown [@rmarkdown2023].
-R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements) and executable computer code (R, Python, or other languages) in a single document [@xie2018r].
+R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements), and executable computer code (R, Python, or other languages) in a single document [@xie2018r].
 Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
@@
 As another example, (R) Markdown also forms the basis of _executable research articles_ [@tsang2020welcome], a browser-based form of research articles that allows authors to enrich their publications with interactive elements, such as figures or results that dynamically change based on readers' input.
-A particular advantage of using R Markdown for creating research reports is that it helps avoiding common threats to reproducibility of statistical results, first and foremost, copy&paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
-further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023] or preregistrations [@peikert2021samething], all using the same Markdown approach.
+A particular advantage of using R Markdown for creating research reports is that it helps to avoid common threats to reproducibility of statistical results, first and foremost, copy-and-paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
+further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023], or preregistrations [@peikert2021samething], all using the same Markdown approach.
 
-In sum, using R Markdown for the creation of scientific reports including statistical computations comes with various advantages, particularly avoiding copy & paste errors as a threat to reproducibility.
+In sum, using R Markdown for the creation of scientific reports including statistical computations comes with various advantages, particularly avoiding copy-and-paste errors as a threat to reproducibility.
 However, there are still various other sources of errors that are not eliminated solely by using R Markdown.
@@
 For another example, bugfixes or changes in package defaults across different package versions may go unnoticed and lead to different results.
-Therefore, it is essential to monitor changes in computational results within R Markdown documents throughout the document's entire lifecycle -- from the initial draft, through (peer-)review, to publication and long-term storage.
+Therefore, it is essential to monitor changes in computational results within R Markdown documents throughout the document's entire lifecycle -- from the initial draft, through (peer) review, to publication and long-term storage.
 Currently, there is no straightforward, automated method for documenting and verifying reproducibility.
@@
 We begin with a straightforward tutorial demonstrating a typical use case for the package. 
-Following this, we delve into more detailed technical background information and exokire various ways to customize the package's behavior.
-While a deep understanding of the technical details isn't necessary for effective use, gaining insight into these aspects can enhance the user experience and help readers appreciate the challenges of ensuring reproducibility.
+We then delve into more detailed technical background information and explain various ways to customize the package's behavior.
+While a deep understanding of the technical details is not necessary for effective use, gaining insight into these aspects can enhance the user experience and help readers appreciate the challenges of ensuring reproducibility.
 
@@
 Successes and failures of the reproduction attempts are displayed in a reproducibility report for each chunk.
-Customizations or suppression of these reports are possible as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a Markdown source file] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+Customizations or suppression of these reports are also possible, as we will demonstrate later.
+A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a Markdown source file.] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at which stages during the data analysis.
-This allows for forensic examination of non-reproducibility; for example, finding out whether the non-reproducibility is due to changes in the data file, due to the data cleaning or the estimation of a statistical model.
+This allows for forensic examination of non-reproducibility; for example, finding out whether the non-reproducibility is due to changes in the data file, due to the data cleaning, or to the estimation of a statistical model.
 Using `reproducibleRchunks`, users can separately assess the reproducibility of each stage of a given statistical data analysis (or any other computation).
 Also, not all elements of a research asset may be required to reproduce (such as the current date displayed in a presentation of results).
-In the following, we will briefly demonstrate how reproduction will be checked and how successful and failed attempts are displayed by the package.
+In the following, we will briefly demonstrate how reproduction is checked and how successful and failed attempts are displayed by the package.
 We will go into details of the mechanics of the package, and discuss potential cases of non-reproducibility and how the package will support researchers in handling these cases.
@@
 
-The core functionality of the package is to provide  `reproducibleR` chunks, which are code chunks in R Markdown documents that can be used like regular `R`-chunks (including name labels and the usual options regulating output parameters) but offer reproducibility checks in addition.
-To use these new code chunks, the package must be loaded^[If an R Markdown file is knitted and the `reproducibleRchunks` package was not loaded, this error is shown: Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'`] in the first regular `R` code chunk in the document using
+The core functionality of the package is to provide  `reproducibleR` chunks, which are code chunks in R Markdown documents that can be used like regular `R` chunks (including name labels and the usual options regulating output parameters) but offer reproducibility checks in addition.
+To use these new code chunks, the package must be loaded^[If an R Markdown file is knitted and the `reproducibleRchunks` package was not loaded, this error is shown: Error `In get_engine(options$engine) :  Unknown language engine 'reproducibleR'`.] in the first regular `R` code chunk in the document using:
 
@@
 On load, the package registers a new type of code chunk called `reproducibleR`.
-Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_.
-The entire tutorial file can be found in the R package in sub directory "inst/examples_simple_demo" or can be downloaded from \url{https://github.com/brandmaier/reproducibleRchunks/tree/main/inst/examples/simple_demo}.
+Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file in which there is a `reproducibleR` chunk with the name _addition_.
+The entire tutorial file can be found in the R package in subdirectory "inst/examples_simple_demo" or can be downloaded from \url{https://github.com/brandmaier/reproducibleRchunks/tree/main/inst/examples/simple_demo}.
 In the chunk, a new variable `my_sum` is declared and defined to be the sum of some variable `x` plus one. 
@@
 
-```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight)"}
+```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight)."}
 knitr::include_graphics("img/rstudio-screenshot-marker2.png")
@@
 
-In this example, we assume that the R Markdown document is eventually converted into a HTML-report (but all other pandoc-supported formats are supported as well, such as PDF or Word documents).
-During initial document conversion, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file (with prefix _.repro_), which contains all information for further reproducibility verification tests.
+In this example, we assume that the R Markdown document will eventually be converted into a HTML-report (but all other pandoc-supported formats are possible as well, such as PDF or Word documents).
+During initial document conversion, the R code in the `reproducibleR`-chunk is executed and the value of `my_sum` is stored in a separate file (with the prefix _.repro_) that contains all information for further reproducibility verification tests.
 A reproducibility report is shown in the generated document below the code chunk indicating information about all variables of the given chunk.
-Figure \@ref(fig:rstudioscreen2) show a snippet from the HTML report generated during the initial creation of the document.
+Figure \@ref(fig:rstudioscreen2) shows a snippet from the HTML report generated during the initial creation of the document.
 
@@
 
-```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is re-generated and a result is successfully reproduced."}
+```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is re-generated and the result is successfully reproduced."}
 knitr::include_graphics("img/generation-step2.png")
@@
 
-If the document is reproduced and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
-In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
+If the document is reproduced and the computational result differs from the original result, the failure to reproduce is noted in the reproducibility report.
+In a variation of the previous example, we assume someone changed the value of `x`, the variable `my_sum` would change, and a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
 
-```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a result failed.", fig.pos="b"}
+```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a result has failed.", fig.pos="b"}
 knitr::include_graphics("img/generation-step2-failed.png")
@@
 For example, when providing Open Code on open platforms, the repository should contain both the Markdown files as well as the metadata files.
-Note that the metadata file format is particularly suitable to be used with version control systems [see @peikert2021reproducible].
+Note that the metadata file format is particularly suitable for use with version control systems [see @peikert2021reproducible].
 If users intentionally modify R code later, e.g., because they want to fix an error, then the metadata files would need to be deleted and recreated.
-Version control can help users in documenting such changes.
+Version control can help users document such changes.
 
@@
 
-In the following, we describe how the package stores, retrieves and compares reproducibility information to verify reproduction attempts.
+In the following, we describe how the package stores, retrieves, and compares reproducibility information to verify reproduction attempts.
 First of all, the package executes `reproducibleR` code chunks just like regular R code chunks, meaning there are no additional restrictions on what can be computed and tested for reproducibility.
@@
 Here is an example of how the contents of a single variable called `numbers` is stored in JSON format.
-Here, the variable content is a vector of five random draws from the number one to ten: `r numbers` generated by the following command:
+In this case, the variable content is a vector of five random draws from the number one to ten: `r numbers` generated by the following command:
 
@@
 The JSON format has a clear, formal structure that facilitates parsing of information for both computers and humans.
-This accessibility can ease forensics if reproduction issues arise.
+This accessibility can make forensics easier if reproduction issues arise.
 To further aid forensic investigation, the package also stores information about the code syntax used to generate an object.
-While JSON can represent arbitrarily complex R objects (such as an entire regression model), we opted to not store all computational results by default but only their "fingerprints".
+While JSON can represent arbitrarily complex R objects (such as an entire regression model), we opted to not store all computational results by default but only their "fingerprints."
 Using the fingerprint, we can determine if an object has changed without storing the object itself.
@@
 ```{r jsonschema, fig.cap="A JSON file stores fingerprints of all newly declared variables in a reproducibleR-chunk as well as fingerprints of the code chunk syntax.
-This image by Andreas Brandmaier and Anke Schneider is licensed under [CC BY 4.0]( https://creativecommons.org/licenses/by/4.0/)"}
+This image by Andreas Brandmaier and Anke Schneider is licensed under CC BY 4.0 ( https://creativecommons.org/licenses/by/4.0/)."}
 knitr::include_graphics("img/schema-json-fingerprints_v4.png")
@@
 Once a document is (re-)generated with metadata present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects that are amenable for reproducibility checks.
-Identicality of original and reproduced results is checked with the `all.equal()` function from R's `base` package, which according to the R documentation is a "is a utility to compare R objects x and y testing ‘near equality’.
-If they are different, comparison is still made to some extent, and a report of the differences is returned." (see `?all.equal` R documentation).
+Identicality of the original and reproduced results is checked with the `all.equal()` function from R's `base` package.
+According to the R documentation, it is a "is a utility to compare R objects x and y testing ‘near equality’.
+If they are different, comparison is still made to some extent, and a report of the differences is returned." [see `?all.equal` R documentation, @R2024].
 
@@
 In principle, any R object is suitable for reproducibility testing no matter how complex.
-For creating fingerprints, the package uses R's serialization methods, which take any R object as input and prepares it for storage.
+For creating fingerprints, the package uses R's serialization methods that take any R object as input and prepare it for storage.
 That is, all of the following variables (`x` of class _integer_, `y` of class _character_, `z` of class _lm_, and `lst` of class _list_) are valid examples for the automated reproducibility checks:
@@
 In the following, we provide a practical guide on how to use reproducible R code chunks in Markdown documents.
-First, to install the package, please install the latest version from CRAN^[The latest version can be found in the package repository: https://github.com/brandmaier/reproducibleRchunks]:
+First, to install the package, please install the latest version from CRAN^[The latest version can be found in the package repository: https://github.com/brandmaier/reproducibleRchunks.]:
 
@@
 Now, we define a reproducible code chunk by setting the code chunk language to `reproducibleR`. 
-Once this document is rendered the first time, a data file will be created, which stores all reproducible results, which are computed in this code block. 
+Once this document is rendered for the first time, a data file will be created that stores all reproducible results computed in this code block. 
 The name of the file will contain the label name (in this example `helloworld`). 
 The following code block contains two reproducible results stored in variables `x` and `y`. 
-Those computations can be based on results computed in previous chunks but those variables will not per se be subject to the reproducibility tests.
+Those computations can be based on results computed in previous chunks but those variables will not be subject to the reproducibility tests _per se_.
 If the R markdown document is rendered a second time, the computational results are recomputed and compared against the original results. 
-For each result, success or failure of reproduction is reported separately.
+For each result, reproduction success or failureis reported separately.
 
@@
 Rendering this markdown twice in a row on the same computer should result in a message of successful replication. 
-To break reproducibility, remove or comment out the `set.seed()` command, which sets the random number generator in a reproducible state, and render the document again. 
+To break reproducibility, remove, or comment out the `set.seed()` command that sets the random number generator in a reproducible state, and render the document again. 
 Now, you should obtain an error message indicating that the result of `x` could not be reproduced while `y` still reproduces as it did not depend on the random number generator.
 
-In practice, we expect that reproducible R chunks will be used both in cases, in which the reproduction report should be displayed (e.g., when developing a data analysis script or reproducing a former one) and those where it should not (e.g., when rendering a presentation or a manuscript for submission to a journal).
+In practice, we expect that reproducible R chunks will be used both in cases in which the reproduction report should be displayed (e.g., when developing a data analysis script or reproducing a former one) and those in which it should not (e.g., when rendering a presentation or a manuscript for submission to a journal).
 By default, reproduction report statements are produced for each chunk but this default can be changed via the code chunk argument `report=FALSE`.

### 2025-04-03 manuscript.Rmd
final set of edits suggested by JD

@@
 `reproducibleRchunks.digits`
-: This is the number of digits for rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $10$.
+: This is the number of digits for the rounding of numbers and controls the numeric precision of the reproducibility checks. By default, this is $10$.
 
 `reproducibleRchunks.filetype`
-: This is the type of data storage. Currently, we only support the 'json' format; however, this leaves room for other formats to be supported in future.
+: This is the type of data storage. Currently, we only support the JSON format; however, this leaves room for other formats to be supported in future. By default, this is 'json'.
 
@@
 `reproducibleRchunks.hashing_algorithm`
-: This is the hashing algorithm, which is used to generate fingerprints of variable contents. By default, the algorithms offered by the `digest` package are used.
+: This is the hashing algorithm that is used to generate fingerprints of variable contents. By default, the algorithms offered by the `digest` package are used.
 
 `reproducibleRchunks.templates`
-: A list with keys corresponding to pandoc output formats and values to templates for formatting the reproducibility reports. See below for more details.
+: A list with keys corresponding to pandoc output formats and values corresponding to templates for formatting the reproducibility reports. See below for more details.
 
@@
 
-There are various hashing functions available, which have different features.
+There are various hashing functions with different features available.
 They generally differ along three dimensions: speed, chance of collisions, and security.
@@
 In this application, the choice of algorithm is not overly crucial.
-Supported algorithm [provided by the digest package, @digest2022] are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash and blake3. 
-To reduce chance of collisions, the package defaults to `sha256` which results in fingerprints of 64 hexadecimal characters, which corresponds to a 256-bit fingerprint but comes in cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers takes `r mean_time_s` seconds on our standard computers).
+Supported algorithm [provided by the digest package, @digest2022] are sha1, crc32, sha256, sha512, xxhash32, xxhash64, murmur32, spookyhash, and blake3. 
+To reduce chance of collisions, the package defaults to `sha256`, which results in fingerprints of 64 hexadecimal characters that corresponds to a 256-bit fingerprint, but comes at the cost of some speed (`sha256` of a vector of `r numbers_in_simulation` numbers still takes only `r mean_time_s` seconds on our standard computers).
 We believe that in most cases, the speed factor will be irrelevant but it could become an issue when very large objects (e.g., entire neuroimaging data sets) are fingerprinted.
-In that case, users either have to use a faster fingerprinting algorithm or forego fingerprinting raw data and instead fingerprint intermediate or final results only.
+In that case, users either have to use a faster fingerprinting algorithm or forego the fingerprinting of raw data and fingerprint intermediate or final results only instead.
 
@@
 Note that these options can be chosen differently for each chunk.
-That is, it is possible, to use fingerprints for storing results of one chunk and plain data storage for results of another chunk.
+That is, it is possible to use fingerprints to store results of one chunk and plain data storage for results of another chunk.
 
@@
 Last, users of this package can customize the appearance of the reproducibility reports. 
-This can be done by either using convenient template layouts, which can be tailored to each specific (pandoc) output format, that is, for example a template for HTML file and a different LaTeX template for PDF files.
-Or, users can obtain a summary of the reproduction status of every variable in every chunk. 
-This information is returned by function `get_reproducibility_summary()` which returns a `data.frame` with three columns. 
+This can be done by either using convenient template layouts, which can be tailored to each specific (pandoc) output format, for example a template for HTML files and a different LaTeX template for PDF files.
+Alternatively, users can obtain a summary of the reproduction status of every variable in every chunk. 
+This information is provided by the function `get_reproducibility_summary()`, which returns a `data.frame` with three columns. 
 The first column contains the name of the code chunk, the second column contains a variable name, and the third column contains a boolean variable representing whether the reproduction was a success. 
@@
 The option `reproducibleRchunks.templates` stores the default templates that are used for displaying the report information. 
-It is a list of key-value pairs where the key is the final pandoc output format (typically either html, pdf or docx) and the value is a string containing formatting information.
-If html output is chosen, it can contain any valid HTML/CSS code, if the output is pdf, it can contain any valid LaTeX code.
+It is a list of key-value pairs where the key is the final pandoc output format (typically either 'html', 'pdf', or 'docx') and the value is a string containing formatting information.
+If 'html' output is chosen, it can contain any valid HTML/CSS code, if the output is 'pdf', it can contain any valid LaTeX code.
 These formatting instructions can contain two placeholders, which the package will replace with the title of the report (\${title}) and the content of the report (\${content}).
@@
 
-As mentioned above, if users wish to entirely suppress the default reproducibility reports, they can use the chunk argument `report=FALSE` as part of the code chunk options to suppress the display. 
+As mentioned above, if users wish to entirely suppress the default reproducibility reports, they can use the chunk argument `report=FALSE` as part of the code chunk options to do so. 
 Note that reproduction is still attempted and reproducibility information is still accessible through the function `get_reproducibility_summary()`.
 The other standard chunk options can be used with `reproducibleR`-code chunks as usual. 
-In particular, `eval=FALSE` suppressed execution of the code, `echo=FALSE` suppressed that the code is shown and so forth.
+In particular, `eval=FALSE` suppresses execution of the code, `echo=FALSE` suppresses that the code is shown and so forth.
 
@@
 Then, all variables created in the process of a given data analysis will be tested for reproducibility.
-However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all statistical results that researchers find meaningful will perfectly reproduce.
+However, we advise against this procedure as it is likely that some variables will not exactly reproduce even though all statistical results that researchers find meaningful will reproduce perfectly.
 
@@
 
-Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error.
-For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
-Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables.
-Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
-For illustration, here is the content of the `cpuTime` variable and the `timestamp` variable for the aforementioned factor model, which was estimated when this manuscript was generated:
+To explain: Statistical models may often store internal information that is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error.
+For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time that elapsed when fitting the model. 
+Below is an example of the contents of a simple confirmatory factor model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables.
+Among information relevant to check reproducibility of parameter estimates and model fit, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
+For illustration, we show the content of the `cpuTime` variable and the `timestamp` variable for the aforementioned factor model, which was estimated when this manuscript was generated:
 
@@
 
-Thus, we generally recommend a checkpoint approach, in which reproducibility tests are used only at selected, meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size point estimates, goodness-of-fit indices, standard error estimates, confidence intervals, test statistics, p values, Bayes factors, etc.).
-Specifically, we recommend to use at least the following checkpoints:
+Therefore, we generally recommend a checkpoint approach in which reproducibility tests are used only at selected, meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (e.g., effect size point estimates, goodness-of-fit indices, standard error estimates, confidence intervals, test statistics, p values, Bayes factors, etc.).
+Specifically, we recommend use of at least the following checkpoints:
 
 - Data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
-- Preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
+- Preprocessing: check whether the preprocessed data (that is, data after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
 - Results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
@@
 A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as parameter estimates, effect size estimates, confidence interval limits, p values, fit indices, or Bayes factors.
-To this end, several packages support the generic function `coef()`, which can be used, for example, to extract point estimates from linear regression models.
-Other packages offer their own accessor function such as `parTable()` for structural equation models in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
-The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with OpenMx models), it may again contain timing information, which will not exactly reproduce.
-Last, note that sometimes there may small deviations of exact reproducibility, which can be tolerated.
-For example, when comparing reproducibility across machines that work with different numerical precision (32bit vs 64bit), numeric representations may count as reproducible if they are identical up to some numerical precision.
+To this end, several packages support the generic function `coef()` that can be used, for example, to extract point estimates from linear regression models.
+Other packages offer their own accessor function such as `parTable()` for structural equation models in `lavaan` or `omxGetParameters()` in the aforementioned `OpenMx` model.
+The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with `OpenMx` models), it may again contain timing information, which will not exactly reproduce.
+Note that small deviations of exact reproducibility that can be tolerated may occur.
+For example, when comparing reproducibility across machines that work with different numerical precision (e.g., 32 bit vs 64 bit precision), numeric representations may count as reproducible if they are identical up to some numerical precision.
 To avoid such problems, `reproducibleRchunks` rounds numeric values to a given precision (by default, ten digits).
-Lastly, users may wish to test identity of reproduction up to a certain precision less than the package default.
-This could be relevant for algorithms that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar (even though, their perfect reproducibility should usually be guarantueed by setting a random seed).
+Finally, users may wish to test identity of reproduction up to a certain precision lower than the package default.
+This could be relevant for algorithms that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators, and similar [even though their perfect reproducibility should usually be guaranteed by setting a random seed, but see @peikert2021reproducible].
 In this case, approximate reproducibility tests could be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
-Adherents of the tidyverse approach [@wickham2019welcome] are well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package.
-These functions convert parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easily accessible format known as _tibble_.
+Adherents of the tidyverse approach [@wickham2019welcome] are well advised to use functions `tidy()` and `glance()` from the `broom` package [@broom].
+These functions convert parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan, smooth.spline, survfit, and others) into a consistent and easily accessible format known as _tibble_.
 This format is particularly suitable for fingerprinting using our proposed approach.
@@
 
-As a rule of thumb, we recommend to fingerprint the results of all parameter estimates from the `tidy()` function and all model fit statistics from the `glance()` function in a `reproducibleR` chunk.
+As a rule of thumb, we recommend fingerprinting the results of all parameter estimates from the `tidy()` function and all model fit statistics from the `glance()` function in a `reproducibleR` chunk.
 
@@
 Dynamic document generation using R Markdown (e.g., reports, posters, or presentations) is an important building block for reproducible research. 
-However, a simple way to automatically verify reproducibility of dynamically generated content within such R Markdown documents is still missing. 
-We developed an R package that enables automatic reproducibility testing of R code chunks within R Markdown documents while requiring essentially no changes to users' workflows.
-This is achieved by storing meta-data about computational results, which can later be used to verify reproducibility.
+However, a simple way to automatically verify reproducibility of dynamically generated content within such R Markdown documents was lacking. 
+We have developed an R package that enables automatic reproducibility testing of R code chunks within R Markdown documents while requiring essentially no changes to users' workflows.
+This is achieved by storing metadata about computational results that can be used later to verify reproducibility.
 Templates allow for customization of the appearance of reproducibility reports; further, the package allows users to generate entirely customized reproducibility reports either within a given R Markdown document or as separate files.
@@
 Adopting reproducibleR chunks in a workflow offers several benefits.
-First of all, it allows to verify reproducibility manually whenever an R Markdown document is rendered. 
+First of all, it allows manual verification of reproducibility whenever an R Markdown document is rendered. 
 This may be particularly relevant when collaboratively working on scientific reports. 
@@
 On collaborative platforms like GitHub, reproducibility could be automatically tested with a GitHub action whenever the repository is modified. 
-It's conceivable that journals could use a similar mechanism to check the reproducibility of a paper upon submission.
+It is conceivable that journals could use a similar mechanism to check the reproducibility of a paper upon submission.
 
@@
 Regular R scripts are not amenable to our proposed approach.
-Recently, Quarto [@quarto2024] was proposed as a next-generation Markdown publishing system and it has been gaining popularity. 
+Recently, Quarto [@quarto2024] was proposed as a next-generation Markdown publishing system and its popularity has grown. 
 However, our proposed approach is not yet compatible with Quarto.
@@
 
-In the following, we briefly discuss a few future situations, in which threats to reproducibility or reproducibility checking happened and explain how these situations can be dealt with:
+In the following, we briefly discuss a few future scenarios in which threats to reproducibility or reproducibility checking may occur and explain how these situations can be dealt with:
 
@@
 Later, a reproduction attempt fails using the exact same R Markdown file executed on a different computer.
-This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021samething].
+This is a classic case of non-reproducibility that is often due to changes in the software packages and the R version the computations in R Markdown rely upon [@epskamp2019reproducibility; @peikert2021samething].
 Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility.
-To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
+To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac].
 - Some computations are executed and their original results are stored in a JSON file as planned.
@@
 In developing this package, we adhere to three major aspects of rigor in scientific software development [@brandmaier2024commentary]. 
-First, the package comes with a variety of formal tests (based on the `testthat` package, @testthat2011) to test correct functioning of the package.
+First, the package comes with a variety of formal tests [based on the `testthat` package, @testthat2011] to test correct functioning of the package.
 Second, we provide documentation in form of this manuscript and an online documentation ([https://github.com/brandmaier/reproducibleRchunks](https://github.com/brandmaier/reproducibleRchunks)). 
-Third, bug reports and feature requests can be submitted through our github project website.
+Third, bug reports and feature requests can be submitted through our GitHub project website.
 
@@
 
-Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatically testing and verifying reproducibility.
-For example, it would allow for reproducibility checks in peer review as suggested by @cruewell2023 could be automated using this package.
+Again, we emphasize that this package is not meant to ensure reproducibility but allows for automatic testing and verification of reproducibility.
+For example, it would allow for reproducibility checks in peer review as suggested by @cruewell2023 which could be automated using this package.
 According to the authors' convention, the current approach relies on testing whether results are _exactly reproducible_, which is consistent with the idea of awarding an Open Data badge.
-If there was a demand, the package could be extended to support the more fine-grained judgements suggested by @cruewell2023, such as _essentially reproducible_ (minor deviations in the decimals), _partially reproducible_ (minor deviations but the results were mostly numerically consistent),or _mostly not reproducible_ (major deviations).
+If there was a demand, the package could be extended to support the more fine-grained judgements suggested by @cruewell2023, such as _essentially reproducible_ (minor deviations in the decimals), _partially reproducible_ (minor deviations, but the results were mostly numerically consistent), or _mostly not reproducible_ (major deviations).
 Note, however, that the consistent use of R Markdown already eliminates some sources of irreproducibility, such as copy-and-paste errors [@peikert2021samething].
-Even though, a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various situations, in which such approaches (including our own) failed.
-First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that execute code under identical conditions on different machines (or the same machine at different time points, e.g., before and after upgrade of R or some or all packages used).
-On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., this happened when Apple switched to their own processor brand) or because a user does not have admin privileges for installing Docker in the first place.
-Further, some approaches rely on service providers such as Microsoft's MRAN archive, which a while ago was unexpectedly terminated. 
-These different weak points now and then lead to situations, in which users will try to locally reproduce historic computational results and they may want to ensure that these reproductions were succesful.
+Even though a variety of approaches have been suggested to ensure reproducibility of computations in R Markdown documents [@peikert2021samething; @renv2022; @chan2023rang; @nagraj2023pracpac], we ourselves have encountered various scenarios in which such approaches (including our own) failed.
+First and foremost, many approaches rely on further software packages such as Docker, which essentially provide virtual environments that execute code under identical conditions on different machines (or the same machine at different time points, e.g., before and after an upgrade of R or of some or all packages used).
+On some machines, Docker may simply be not available, either because it is not (yet) available on a certain newly introduced hardware (e.g., this happened when Apple switched to their own processor brand) or because a user does not have admin privileges to install Docker in the first place.
+Further, some approaches rely on service providers such as Microsoft's MRAN archive, which  was unexpectedly terminated a while ago. 
+Now and then, these different weak points lead to situations, in which users will try to locally reproduce historic computational results and may want to ensure that these reproductions were succesful.
 With our package, the success of such reproduction attempts becomes easily and formally testable.
 
-We hope the suggested approach helps raising awareness for the importance of reproduciblity and increasing the visibility of potential non-reproducibility issues in R code, eventually increasing quality and credibility of digital research assets.
+We hope the suggested approach helps to raise awareness for the importance of reproduciblity and increasing the visibility of potential non-reproducibility issues in R code, eventually increasing quality and credibility of digital research assets.
 \newpage

### 2025-04-07 manuscript.Rmd
differentiate reproduction and rerendering closes #11

@@
 R Markdown is a simple language, which allows mixing natural language, simple formatting instructions (e.g., what is a headline, what should appear in bold face, or what are list elements), and executable computer code (R, Python, or other languages) in a single document [@xie2018r].
-Whenever an R Markdown document is reproduced, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
+Whenever an R Markdown document is rendered, all of its computer code chunks are executed, which allows for the generation of dynamic content in the document.
 This is particularly interesting for quantitative research reports, as  statistical results and contents of tables and figures can be dynamically generated.
@@
 A particular advantage of using R Markdown for creating research reports is that it helps to avoid common threats to reproducibility of statistical results, first and foremost, copy-and-paste errors, which happen when results of a regular R script and a scientific report (e.g., a reported statistic or table in a Word document) mismatch by mistake [@peikert2021reproducible];
-further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023], or preregistrations [@peikert2021samething], all using the same Markdown approach.
+further, R Markdown is versatile enough to generate a variety of scientific assets beyond manuscripts, including posters [@posterdown2019], presentations [@xaringan2022], resumés [@vitae2023], or preregistrations [@peikert2021samething], all using the same R Markdown approach.
 
@@
 For another example, bugfixes or changes in package defaults across different package versions may go unnoticed and lead to different results.
-Therefore, it is essential to monitor changes in computational results within R Markdown documents throughout the document's entire lifecycle -- from the initial draft, through (peer) review, to publication and long-term storage.
+Therefore, it is essential to monitor changes in computational results within R Markdown documents throughout the document's entire lifecycle --- from the initial draft, through (peer) review, to publication and long-term storage.
 Currently, there is no straightforward, automated method for documenting and verifying reproducibility.
@@
 Furthermore, information about the code chunk itself (that is, the code syntax) is also stored, such that later non-reproduction can be traced back to either changes of the code chunk or changes of the computational environment.
-Once a document is reproduced, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
+Once a document is rendered again, all computational results generated in the `reproducibleR` chunks are automatically tested for reproducibility.
 Successes and failures of the reproduction attempts are displayed in a reproducibility report for each chunk.
 Customizations or suppression of these reports are also possible, as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a Markdown source file.] Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a R Markdown source file.] R Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at which stages during the data analysis.
@@
 
-Once the document is reproduced, that is, generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
+Once the document is reredered, that is, generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
 The reproducibility report will then detail the success of each reproduction attempt.
@@
 
-If the document is reproduced and the computational result differs from the original result, the failure to reproduce is noted in the reproducibility report.
+If the document is reredered and the computational result differs from the original result, the failure to reproduce is noted in the reproducibility report.
 In a variation of the previous example, we assume someone changed the value of `x`, the variable `my_sum` would change, and a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).
@@
 Once a document is (re-)generated with metadata present (i.e., reproducibleRchunks with matching JSON files), the package will compare all objects that are amenable for reproducibility checks.
-Identicality of the original and reproduced results is checked with the `all.equal()` function from R's `base` package.
+Identicality of the original and regenerated results is checked with the `all.equal()` function from R's `base` package.
 According to the R documentation, it is a "is a utility to compare R objects x and y testing ‘near equality’.

### 2025-04-10 manuscript.Rmd
Update manuscript.Rmd

@@
 Customizations or suppression of these reports are also possible, as we will demonstrate later.
-A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from a R Markdown source file.] R Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
+A simple way to assess reproducibility of computational results would be a bitwise comparison of the entire knitted^[This is R Markdown-speak for generating a publishable document from an R Markdown source file.] R Markdown document (i.e., the resulting HTML, PDF or DOCX file) to a previous version.
 Yet, a particular advantage of using `reproducibleR` chunks is that users have a fine-grained control over what should be tested for reproducibility and at which stages during the data analysis.

### 2025-04-10 manuscript.Rmd
Update manuscript.Rmd

@@
 
-Once the document is reredered, that is, generated a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
+Once the document is rendered a second time, and the reproducibility metadata file exists, the computations in the code chunk are rerun, and their results are compared to the stored reproducibility information for each variable.
 The reproducibility report will then detail the success of each reproduction attempt.

### 2025-04-10 manuscript.Rmd
Update manuscript.Rmd

@@
 
-If the document is reredered and the computational result differs from the original result, the failure to reproduce is noted in the reproducibility report.
+If the document is rendered again and the computational result differs from the original result, the failure to reproduce is noted in the reproducibility report.
 In a variation of the previous example, we assume someone changed the value of `x`, the variable `my_sum` would change, and a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).

### 2025-06-03 manuscript.Rmd
fix missing space added ethics statement

@@
 The manuscript source code is available at https://github.com/brandmaier/reproducibleRchunks-manuscript.
-We thank Julia Delius forher helpful assistance in language and style editing.
+We thank Julia Delius for her helpful assistance in language and style editing.
 We also thank Leonie Hagitte for providing comments on an earlier version of the manuscript.
@@
 #output            : papaja::apa6_docx
+keeptex           : yes
 header-includes:
@@
 If the R markdown document is rendered a second time, the computational results are recomputed and compared against the original results. 
-For each result, reproduction success or failureis reported separately.
+For each result, reproduction success or failure is reported separately.
 
@@
 Adopting reproducibleR chunks in a workflow offers several benefits.
-First of all, it allows manual verification of reproducibility whenever an R Markdown document is rendered. 
-This may be particularly relevant when collaboratively working on scientific reports. 
+First of all, it allows manual verification of reproducibility whenever an R Markdown document is rendered. This may be particularly relevant when collaboratively working on scientific reports. 
 Adopting the proposed approach ensures that the same results are generated across all computers of a team of researchers. 
@@
 
+# Ethics statement
+
+This study did not involve testing of human participants.
+
 # Data accessibility statement
